---
title: "DE_Milo_Tissue_Myeloid Network expansion"
author: "Bidossessi Hounkpe"
date: "05/12/2025"
format: html
editor: visual
---

```{r}
rm(list=ls())
```

## Disease summary

```{r}
pacman::p_load(
  OmnipathR,
  igraph,
  ggraph,
  graphlayouts,
  ggforce,
  tidygraph,
  httr,
  SingleCellExperiment,
  scran,
  scuttle,
  ggprism,
  ComplexHeatmap,
  pheatmap,
  circlize,
  reticulate,
  biomaRt,
  ghql,
  jsonlite,
  tidyverse,
  RColorBrewer
)

#Use conda env for reticulate
use_condaenv("ScGwas", required = TRUE)

basepath <- "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/scrnaseq-autoimmune-atlas/notebooks/4-R/"
#### Windows
source(str_glue(basepath, "colors.R"))

# Source helper functions

source(str_glue(basepath, "helper_fun.R"))

# Network expansion 

source(str_glue(basepath, "openTarget.R"))

source(str_glue(basepath, "chembl_new.R"))

```

```{python}
import os,sys
import numpy as np
import pandas as pd
#import drug2cell as d2c
#import scanpy as sc
import sctk
#import gc
#sys.path.append("/mnt/c/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

sys.path.append("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

from inflapy.immunoFunc import *
import gseapy as gp
```

```{r}

figurePath = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/All_Tissue_or_Nonimmune/Subset_Reintegrated_full/Figures/"
```

```{r}
efo_disease_genetic = read.csv("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/GWAS_assoc/Efo_disease_genesAssoc.csv")

# EFO codes for autoimmune diseases
efo_codes <- c(
  IBD = "EFO_0003767",
  Rheumatoid_Arthritis = "EFO_0000685",
  JIA = "EFO_0002609",
  Sjogren_Syndrome = "EFO_0000699",
  spondyloarthropathy = "EFO_0000706",
  Hashimoto_Thyroiditis = "EFO_0003779",
  SLE = "MONDO_0007915",
  Psoriatic_Arthritis = "EFO_0003778",
  Atopic_Dermatitis = "EFO_0000274",
  Eczematoid_dermatitis = "HP_0000964"
)

# Use with the function
efo_disease_genetic_results <- get_opentargets_overall_association(
  efo_codes = unname(efo_codes),
  score_threshold = 0,
  page_size = 3000
)

efo_disease_genetic_results$is_truth = ifelse(
  efo_disease_genetic_results$genetic_association_score > 0 & !is.na(efo_disease_genetic_results$genetic_association_score),
  1, 0
)

# Use the saved version for reproducibility
efo_disease_genetic_results <- read.csv( "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/scrnaseq-autoimmune-atlas/notebooks/data/efo_disease_genetic_results.csv")

efo_disease_genetic_results <- efo_disease_genetic_results[,-1]

IL1B_macr_up = read.csv("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/DEG/Enriched/Shared/Results_DE_shared_ILB_Mac_RA_IBD.csv")

IL1B_macr_up = IL1B_macr_up[,-1]

colnames(IL1B_macr_up)[1] = "gene_symbol"

IL1B_macr_up = inner_join(IL1B_macr_up, efo_disease_genetic_results, by = "gene_symbol")

IL1B_macr_up <- IL1B_macr_up %>% mutate(
  disease_name_clean = case_when(
    disease_name == "inflammatory bowel disease" ~ "IBD",
    disease_name == "juvenile idiopathic arthritis" ~ "JIA",
    disease_name == "rheumatoid arthritis" ~ "RA",
    disease_name == "atopic eczema" ~ "Atopic Eczema",
    disease_name == "spondyloarthropathy" ~ "Spondyloarthritis",
    disease_name == "systemic lupus erythematosus" ~ "SLE",
    disease_name == "Sjogren syndrome" ~ "Sjögren syndrome",
    disease_name == "Eczematoid dermatitis" ~ "Atopic Eczema",
    TRUE ~ disease_name
    
  )
)

IL1B_macr_up_wide = pivot_wider(unique(IL1B_macr_up[,c(1,14,13)]), names_from = disease_name_clean, values_from = is_truth, values_fill = 0, values_fn = max)

IL1B_macr_up_wide = tibble::column_to_rownames(IL1B_macr_up_wide, var = "gene_symbol")

IL1B_macr_up_wide = IL1B_macr_up_wide[rowSums(IL1B_macr_up_wide) != 0,]

IL1B_macr_up_wide = IL1B_macr_up_wide[, colSums(IL1B_macr_up_wide) != 0]
```

### Figure 6B

```{r fig.width=10, fig.height=29}


heat_m = Heatmap(IL1B_macr_up_wide,
        cluster_columns = F,
        name = "Genetic Association",
        col = structure(c("grey80", "grey25"), names = c(0, 1)),
        width = ncol(IL1B_macr_up_wide) * unit(10, "mm"),
        height = nrow(IL1B_macr_up_wide) * unit(10, "mm"),
        show_row_dend = T,
        row_names_side = "right",
        rect_gp = gpar(col = "white"),
        row_dend_width = unit(3, "cm"),
        row_dend_gp = gpar(lwd =3),
        heatmap_legend_param = list(
        title_gp = gpar(fontsize = 25, fontface = "bold"),
        labels_gp = gpar(fontsize = 25),
        at = c(0, 1), 
        labels = c("NO", "Yes"),
        grid_width = unit(12, "mm"),
        grid_height = unit(12, "mm"),
        title_gap = unit(5, "mm"),
        direction = "horizontal"),
      row_names_gp = gpar(fontsize = 25),
      column_names_gp = gpar(fontsize = 25),
      column_title_gp = gpar(fontsize = 30, fontface = "bold"),
      column_title = expression(IL1B^high ~ macrophages))

#heat_m = ggplotify::as.ggplot(heat_m)
  
draw(heat_m, heatmap_legend_side = "right")

```

```{r}

svg(str_glue("{figurePath}Heatmap_IL1B_Macr_up_Genetic_OPT_Mac_2025.svg"), width = 10, height = 29)
draw(heat_m, heatmap_legend_side = "right")
dev.off()


```

### Data loading and preparation

```{r}

sce_path = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/All_Tissue_or_Nonimmune/Subset_Reintegrated_full/Milo/enriched_depleted_SCE/"

# 

load(str_glue("{sce_path}Pseudo_test_adata_individual_disease_Myel_Tissue_reajusted.RData"))

test_adata_$Cell_type_scANVI = as.character(test_adata_$Cell_type_scANVI)

test_adata_ = test_adata_[, test_adata_$Cell_type_scANVI != "Unknown"]

test_adata_$Cell_type_scANVI = gsub(" ", "_", test_adata_$Cell_type_scANVI) %>% factor()

test_adata_$Cell_type_scANVI = gsub("[:+:]", "_plus", test_adata_$Cell_type_scANVI) %>% factor()

test_adata_$Cell_type_scANVI = gsub("-", "_", test_adata_$Cell_type_scANVI) %>% factor()

test_adata_$donor_id = factor(gsub(" ", "_", test_adata_$donor_id))

test_adata_$donor_id = factor(gsub("-", "_", test_adata_$donor_id))

test_adata_$disease_short = droplevels(test_adata_$disease_short)

test_adata_$Cell_type_scANVI = str_replace(test_adata_$Cell_type_scANVI, "Mo_Mac$", "Mac_CX3CR1")

test_adata_$Cell_type_scANVI = str_replace(test_adata_$Cell_type_scANVI, "Mo_Mac", "Mac")

test_adata_$Cell_type_scANVI = str_replace(test_adata_$Cell_type_scANVI, "Intermediate_macrophage", "IL1B_macrophage")

unique(test_adata_$Cell_type_scANVI)
```

```{r}

test_adata_$tissue = gsub( " ", "_", test_adata_$tissue)

test_adata_$tissue_disease = paste0(test_adata_$tissue, "_", test_adata_$disease_short) %>% factor()


test_adata_$tissue_disease = str_replace(test_adata_$tissue_disease, "Ileum_CD", "ileum_CD") %>% factor()


levels(test_adata_$tissue_disease)


```

```{r}
gene_expressed = psdobkl2expressed(scedata = test_adata_, min.count=10, sample.size.percent = 0.1)

# Download protein-protein interactions

interactions = OmnipathR::import_omnipath_interactions() %>% as_tibble()

# Convert to igraph objects:

OPI_g = interaction_graph(interactions = interactions)
```

## RA and CD signature in intermediate macrophage
### Network expansion

```{r}

IL1B_macr_up_wide_2 = IL1B_macr_up_wide

IL1B_macr_up_wide_2 = IL1B_macr_up_wide_2 %>% mutate(Origin = paste0("Genetic_Assoc_in_", rowSums(IL1B_macr_up_wide_2)),
                             Score = rowSums(IL1B_macr_up_wide_2)/ncol(IL1B_macr_up_wide_2))

IL1B_macr_up_wide_2 = tibble::rownames_to_column(IL1B_macr_up_wide_2, var = "name")

net = disease2net(
  graph = OPI_g,
  poi.meta = IL1B_macr_up_wide_2[,c(1,ncol(IL1B_macr_up_wide_2)-1,ncol(IL1B_macr_up_wide_2))],
  exclude = NULL,
  intermediate = FALSE,
  include = gene_expressed,
  walk.step = 5,
  min.gene.cluster = 5
)
```

```{r}
plot_net = ggraph(net, layout = "fr") +
  geom_edge_link0(width=0.2, colour = "gray90", alpha = 0.5) +
  scale_color_brewer(palette = "Paired") +
  geom_edge_fan(aes(alpha = after_stat(index)), show.legend = FALSE) +
  geom_node_point(aes(color = cwt, size = page.rank)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph()


plot_net

ggsave(str_glue("{figurePath}Tissue_IntermediateMacr_RA_top5_ranked_Modules_2025.svg"), bg='transparent')
```
#### Known drug targets
```{r}

known_drugs_subnet_expansion <- map_df(as_ids(V(net)), function(g) {
  Sys.sleep(0.5)
  tryCatch(get_opentargets_knowndrugs_detailed(g, page_size = 3000), error = function(e) {
    message("  Error: ", e$message)
    NULL
  })
})

known_drugs_subnet_expansion_unique_phase4 <- unique(known_drugs_subnet_expansion[known_drugs_subnet_expansion$status=="Completed" & known_drugs_subnet_expansion$clinical_precedence=="Approved" & !known_drugs_subnet_expansion$indication_is_therapeutic_area &
known_drugs_subnet_expansion$indication_max_phase == 4, c(1,2,5,7,12:16,24:28)]) %>% na.omit()
```


```{r}
diseases <- c(
  "neoplasm",
  "childhood acute lymphoblastic leukemia",
  "Blast Phase Chronic Myelogenous Leukemia, BCR-ABL1 Positive",
  "cancer",
  "lymphoblastic lymphoma",
  "chronic myelogenous leukemia",
  "acute lymphoblastic leukemia",
  "lymphoid leukemia",
  "B-cell acute lymphoblastic leukemia",
  "childhood cancer",
  "dermatofibrosarcoma protuberans",
  "myelodysplastic/myeloproliferative disease",
  "Hypereosinophilic syndrome",
  "gastrointestinal stromal tumor",
  "colorectal neoplasm",
  "colorectal carcinoma",
  "colorectal cancer",
  "colorectal adenocarcinoma",
  "myeloid leukemia",
  "marginal zone B-cell lymphoma",
  "follicular lymphoma",
  "chronic bronchitis",
  "chronic obstructive pulmonary disease",
  "emphysema",
  "myocardial infarction",
  "hypertension",
  "congestive heart failure",
  "heart failure",
  "asthma",
  "Airway obstruction",
  "multiple system atrophy",
  "orthostatic hypotension",
  "open-angle glaucoma",
  "glaucoma",
  "bronchitis",
  "obesity",
  "Nasal congestion",
  "common cold",
  "seasonal allergic rhinitis",
  "hypotension",
  "breast cancer",
  "breast neoplasm",
  "acute myeloid leukemia",
  "Mantle cell lymphoma",
  "chronic lymphocytic leukemia",
  "Waldenstrom macroglobulinemia",
  "macroglobulinemia",
  "alopecia areata",
  "thyroid cancer",
  "papillary thyroid carcinoma",
  "medullary thyroid gland carcinoma",
  "thyroid neoplasm",
  "thyroid carcinoma",
  "acne",
  "breast carcinoma",
  "menopause",
  "Premature ovarian insufficiency",
  "Dyspareunia",
  "Atrophic Vaginitis",
  "primary ovarian insufficiency",
  "postmenopausal osteoporosis",
  "hypogonadism",
  "Atrophy",
  "postmenopausal atrophic vaginitis",
  "osteoporosis",
  "HER2 Positive Breast Carcinoma",
  "infertility",
  "polycystic ovary syndrome",
  "hemorrhage",
  "Menorrhagia",
  "pregnancy",
  "Abnormal retinal morphology",
  "eye disease",
  "bipolar disorder",
  "bipolar I disorder",
  "depressive disorder",
  "major depressive disorder",
  "multiple myeloma",
  "Cutaneous T-cell lymphoma",
  "unspecified peripheral T-cell lymphoma",
  "T-cell non-Hodgkin lymphoma",
  "mature T-cell and NK-cell non-Hodgkin lymphoma",
  "Graves ophthalmopathy",
  "hypothyroidism",
  "growth hormone insensitivity syndrome",
  "Laron syndrome",
  "Growth delay",
  "CINCA syndrome",
  "familial Mediterranean fever",
  "mevalonate kinase deficiency",
  "gout",
  "juvenile idiopathic arthritis",
  "Muckle-Wells syndrome",
  "cryopyrin-associated periodic syndrome",
  "Familial cold urticaria",
  "pericarditis",
  "rheumatoid arthritis",
  "COVID-19",
  "pustulosis palmaris et plantaris",
  "psoriasis",
  "cutaneous melanoma",
  "melanoma",
  "metastatic melanoma",
  "kidney transplant",
  "multiple sclerosis",
  "relapsing-remitting multiple sclerosis",
  "severe acute respiratory syndrome",
  "systemic scleroderma",
  "temporal arteritis",
  "polyarticular juvenile idiopathic arthritis",
  "neuromyelitis optica",
  "dry eye syndrome",
  "ulcerative colitis",
  "Crohn's disease",
  "Myelofibrosis",
  "Vitiligo",
  "primary myelofibrosis",
  "Eczematoid dermatitis",
  "myeloproliferative disorder",
  "graft versus host disease",
  "polycythemia vera",
  "essential thrombocythemia",
  "Splenomegaly",
  "psoriatic arthritis",
  "spotted fever",
  "anthrax infection",
  "brucellosis",
  "Vibrio infectious disease",
  "periodontitis",
  "rosacea",
  "bacterial disease",
  "typhus",
  "infection",
  "pneumonia",
  "ornithosis",
  "gonorrhea",
  "malaria",
  "Tularemia",
  "Friedreich ataxia",
  "bone osteosarcoma",
  "osteosarcoma",
  "neoplasm of mature B-cells",
  "Mastocytosis",
  "systemic mastocytosis",
  "mast-cell leukemia",
  "anemia (phenotype)",
  "anemia",
  "myelodysplastic syndrome",
  "actinic keratosis",
  "autoimmune thrombocytopenic purpura",
  "Thrombocytopenia",
  "Beta-thalassemia",
  "Thalassemia",
  "ankylosing spondylitis",
  "hidradenitis suppurativa",
  "psoriasis vulgaris",
  "psoriasis-related juvenile idiopathic arthritis",
  "uveitis",
  "arthritis",
  "inflammation",
  "spondylitis",
  "spondyloarthropathy"
)

```


```{r}
cancer_keywords <- paste(
  c(
    "cancer",
    "neoplasm",
    "carcinoma",
    "leukemia",
    "lymphoma",
    "myeloma",
    "sarcoma",
    "tumou?r",     # tumor / tumour
    "blastoma",
    "myelofibrosis",
    "myelodysplastic",
    "myeloproliferative",
    "mastocytosis",
    "polycythemia vera",
    "thrombocythemia"
  ),
  collapse = "|"
)

diseases_no_cancer <- diseases[!grepl(cancer_keywords, diseases, ignore.case = TRUE)]

```


```{r}
autoimmune_diseases <- c(
  "alopecia areata",
  "Graves ophthalmopathy",
  "hypothyroidism",
  "juvenile idiopathic arthritis",
  "polyarticular juvenile idiopathic arthritis",
  "rheumatoid arthritis",
  "psoriatic arthritis",
  "psoriasis",
  "psoriasis vulgaris",
  "psoriasis-related juvenile idiopathic arthritis",
  "vitiligo",
  "ulcerative colitis",
  "Crohn's disease",
  "ankylosing spondylitis",
  "uveitis",
  "systemic scleroderma",
  "multiple sclerosis",
  "relapsing-remitting multiple sclerosis",
  "neuromyelitis optica",
  "autoimmune thrombocytopenic purpura",
  "graft versus host disease",
  "systemic mastocytosis",
  "mastocytosis",
  "dry eye syndrome",
  "temporal arteritis",
  "pericarditis",
  "inflammation",

  # Autoinflammatory syndromes (often grouped with autoimmune)
  "familial Mediterranean fever",
  "cryopyrin-associated periodic syndrome",
  "Familial cold urticaria",
  "Muckle-Wells syndrome",
  "CINCA syndrome",
  "mevalonate kinase deficiency"
)

```


##### Phase 4 Drug → Gene → Indication

```{r}
# known_drugs_subnet_RA_CD_unique = known_drugs_subnet_RA_CD %>% 
#   filter(is_approved, status == "Completed", drug_type %in% c("Protein", "Antibody", "Antibody drug conjugate")) %>% 
#   mutate(
#   known_drugs = case_when(is.na(drug_name) ~ "Unknown",
#                           TRUE ~ "Known drugs") 
# )


disease_ids_efo <- c("MONDO_0019751", "MONDO_0007179")  # autoinflammatory syndrome and autoimmune
all_children_autoimmune_autoinflammation <- get_efo_descendants(
  disease_ids_efo, ontology = "mondo")


known_drugs_subnet_expansion_unique_approved <-
  known_drugs_subnet_expansion_unique_phase4 %>%
  select(
    gene,
    drug_name,
    drug_type,
    indication_name,
    indication_max_phase
  ) %>%
  mutate(
    drug_gene = paste(drug_name, gene, sep = "->"),
    value = 1
  ) %>%
  distinct() %>%
  filter(
    toupper(indication_name) %in% toupper(autoimmune_diseases)
  )



known_drugs_subnet_expansion_unique_approved_mat = known_drugs_subnet_expansion_unique_approved %>%
  select(drug_gene, indication_name, value) %>% unique() %>% 
  pivot_wider(names_from = "drug_gene", values_from = "value", values_fill = 0) %>% 
  tibble::column_to_rownames(var = "indication_name")

rownames(known_drugs_subnet_expansion_unique_approved_mat) = gsub(",", "\n", rownames(known_drugs_subnet_expansion_unique_approved_mat))
#known_drugs_subnet_RA_CD_unique = known_drugs_subnet_RA_CD_unique[,c(1,3,5,8)]
```

```{r fig.width=38, fig.height=35}


heat_m_drug_subnet_expansion = Heatmap(t(known_drugs_subnet_expansion_unique_approved_mat),
        cluster_columns = T,
        name = "Indication",
        col = structure(c("grey80", "grey25"), names = c(0, 1)),
        width = nrow(known_drugs_subnet_expansion_unique_approved_mat) * unit(18, "mm"),
        height = ncol(known_drugs_subnet_expansion_unique_approved_mat) * unit(18, "mm"),
        show_row_dend = F,
        show_column_dend = F,
        rect_gp = gpar(col = "white"),
        row_dend_width = unit(3, "cm"),
        row_dend_gp = gpar(lwd =3),
        heatmap_legend_param = list(
        title_gp = gpar(fontsize = 25, fontface = "bold"),
        labels_gp = gpar(fontsize = 25),
        at = c(0, 1), 
        labels = c("NO", "Yes"),
        grid_width = unit(12, "mm"),
        grid_height = unit(12, "mm"),
        title_gap = unit(5, "mm"),
        direction = "horizontal"),
      row_names_gp = gpar(fontsize = 25),
      row_names_side = "left",
      column_names_gp = gpar(fontsize = 25),
      column_title_gp = gpar(fontsize = 30, fontface = "bold"),
      column_names_side = "top", column_title_side = "bottom",
      column_title = "Approved drugs",
      column_names_rot = 45
      )

#heat_m = ggplotify::as.ggplot(heat_m)
  
draw(heat_m_drug_subnet_expansion, heatmap_legend_side = "right")

```

```{r}

svg(str_glue("{figurePath}Heatmap_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_subnet_expansion_2025.svg"), width = 38, height = 35)
draw(heat_m_drug_subnet_expansion, heatmap_legend_side = "right")
dev.off()


```

##### Tripartite plot

```{r fig.width=32, fig.height=30}
known_drugs_subnet_expansion_unique_approved_tripartite = known_drugs_subnet_expansion_unique_approved
known_drugs_subnet_expansion_unique_approved_tripartite$indication_name = str_replace(known_drugs_subnet_expansion_unique_approved_tripartite$indication_name, " Positive", "+") %>%
  str_to_sentence()

known_drugs_subnet_expansion_unique_approved_tripartite$indication_name <- case_when(str_detect(known_drugs_subnet_expansion_unique_approved_tripartite$indication_name, regex("hypertension", ignore_case = TRUE)) ~ "Hypertension", TRUE ~ known_drugs_subnet_expansion_unique_approved_tripartite$indication_name)

create_tripartite_ggplot(
  data = known_drugs_subnet_expansion_unique_approved_tripartite, 
  drug_col = "drug_name", 
  gene_col = "gene", 
  indication_col = "indication_name",
  weight_col = "indication_max_phase",
  color_by = "gene", gene_color = "Set1",
  node_size = 4, label_size = 7.5,
  text_size = 20,
  pathwidth = 2,
  text_face = "plain",
  trial_dot_size = 20,
  size_trial_lim = 0.018,
  size_gene_lim = 0.01,
  trial_dot_color = "grey50",
  title = "Approved drug → Gene → Indication Colored by Drug"
)


```

```{r}
ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_tripartite_subnet_expansion_2025.svg"), width = 32, height = 30, bg = "white")

ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_tripartite_subnet_expansion_2025.png"), width = 32, height = 30, bg = "white")
```

##### Phase 3 Drug → Gene → Indication

```{r fig.width=35, fig.height=80}

known_drugs_subnet_expansion_unique = known_drugs_subnet_expansion %>% mutate(
  known_drugs = case_when(is.na(drug_name) ~ "Unknown",
                          TRUE ~ "Known drugs") 
)

known_drugs_subnet_expansion_unique_phase3 = known_drugs_subnet_expansion_unique %>% 
  select(gene, drug_name, drug_type, indication_name, indication_max_phase, status) %>% 
  filter(indication_max_phase %in% (1:3), status == "Completed") %>% 
  mutate(drug_gene = paste(drug_name, gene, sep = "->")) %>% unique() %>% mutate(value = 1)


known_drugs_subnet_expansion_unique_approved_tripartite_phase3 = known_drugs_subnet_expansion_unique_phase3
known_drugs_subnet_expansion_unique_approved_tripartite_phase3$indication_name = str_replace(known_drugs_subnet_expansion_unique_approved_tripartite_phase3$indication_name, " Positive", "+") %>%
  str_to_sentence()

known_drugs_subnet_expansion_unique_approved_tripartite_phase3$phase_for_indication2 = as.character(known_drugs_subnet_expansion_unique_approved_tripartite_phase3$indication_max_phase)

# create_tripartite_ggplot_2(
#   known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3, 
#   drug_col = "drug_name", 
#   gene_col = "gene", 
#   indication_col = "indication_name",
#   weight_col = "indication_max_phase",
#   color_by = "phase_for_indication2",
#   color_name = "Phase for indication",
#   node_size = 4, label_size = 8.5,
#   custom_color = c("#1f77b4", "#2ca02c", "#d62728"),
#   text_size = 22,
#   pathwidth = 2,
#   text_face = "plain", line_alpha = 0.8,
#   title = "Drug → Gene → Indication - Colored by Drug (phase 1 to 3)"
# )

tripartite_subnet_expansion_ggplot3 <- create_tripartite_ggplot(
  data = known_drugs_subnet_expansion_unique_approved_tripartite_phase3, 
  drug_col = "drug_name", 
  gene_col = "gene", 
  indication_col = "indication_name",
  weight_col = "indication_max_phase",
  color_by = "gene", gene_color = "Set1",
  node_size = 4, label_size = 7.5,
  text_size = 20,
  pathwidth = 2,
  text_face = "plain",
  trial_dot_size = 20,
  trial_dot_color = "grey50",
  title = "Approved drug → Gene → Indication Colored by Drug"
)

tripartite_subnet_expansion_ggplot3

# ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_drugs_phase_1_to_3_tripartite_subnet_expansion_2025.svg"), width = 35, height = 28, bg = "white")
# 
# ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_drugs_phase_1_to_3_tripartite_subnet_expansion_2025.png"), width = 35, height = 28, bg = "white")
```

```{r fig.fullwidth = TRUE, fig.width=6, fig.height=1.2}
drugs_ori_subnet_expansion = known_drugs_subnet_expansion_unique %>%
  filter(status == "Completed" | known_drugs == "Unknown") %>% 
  dplyr::select(gene, known_drugs, clinical_precedence) %>%
  mutate(known_drugs = case_when(known_drugs == "Known drugs" ~ "Yes", TRUE ~ "NO")) %>% 
  #filter(gene != "CD274") %>% # remove PD-L1 (for cancer)
  unique() 
known_drugs_ori_subnet_expansion = drugs_ori_subnet_expansion %>% dplyr::group_by(known_drugs) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count)

barplot4netannot(
  data = known_drugs_ori_subnet_expansion,
  X = "known_drugs",
  Y = "Count",
  label = TRUE,
  ylab = "Number of genes",
  name = "Genes with \n known drug",
  key.size = 0.4,
  colmap = vega_20[c(12,11)],
  label_extra_position = 5,
  flip = TRUE
)

ggsave(str_glue("{figurePath}known_drugs_subnet_expansion_unique_2025.png"), bg='white')
```

```{r fig.height=1.2, fig.width=6}
barplot4netannot(
  data = known_drugs_ori_subnet_expansion,
  X = "known_drugs",
  Y = "Count",
  label = TRUE,
  ylab = "Number of genes",
  name = "Genes with \n known drug",
  key.size = 0.4,
  colmap = vega_20[c(12,11)],
  label_extra_position = 5,
  flip =TRUE
)

ggsave(str_glue("{figurePath}known_drugs_subnet_expansion_unique_2025.png"), bg='white')
```



```{r fig.fullwidth = TRUE, fig.width=3, fig.height=3.5}
# known_drugs_ori_2 = known_drugs_subnet_RA_CD_unique[known_drugs_subnet_RA_CD_unique$known_drugs == "Known drugs",] %>% 
#   filter(status == "Completed") %>% 
#   dplyr::select(gene, clinical_precedence, status, indication_name) %>% 
#   na.omit() %>% unique()# %>% 
known_drugs_ori_2_subnet_expansion = drugs_ori_subnet_expansion %>% na.omit() %>% 
  filter(known_drugs == "Yes") %>% 
  dplyr::group_by(clinical_precedence) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count) %>% 
  dplyr::mutate(clinical_precedence = as.factor(clinical_precedence))

known_drugs_ori_2_subnet_expansion$clinical_precedence = factor(known_drugs_ori_2_subnet_expansion$clinical_precedence,
                                               levels = c("Phase I", "Phase II", "Phase III", "Approved"))

barplot4netannot(
  data = known_drugs_ori_2_subnet_expansion,
  X = "clinical_precedence",
  Y = "Count",
  label = TRUE,
  name = "Clinical Trial \n Precedence",
  key.size = 0.4,
  label_extra_position = 2.5,
  flip = FALSE,
  colmap =  c("#e7e7e7", "#D4B996", "#99847A", "#3A405A")#vega_20
)

ggsave(str_glue("{figurePath}known_drugs_subnet_expansion_Phase_2025.png"), bg='transparent')
```


### Open network in cytoscape

```{r}
library(RCy3)
library(dplyr)
library(igraph)
```

```{r}
cytoscapePing()
```

```{r}

createNetworkFromIgraph(net, "Tissue_IntermediateMacr_RA_CD_from_geneticOPT_5modules_2025")

```

## Known drug targets

```{r}
#library(biomaRt)
#known_drugs_subnet_RA_CD = gene2knowndrugs(geneid = as_ids(V(net)), ensId = FALSE)


known_drugs_subnet_RA_CD <- map_df(unique(IL1B_macr_up_wide_2$name), function(g) {
  Sys.sleep(0.5)
  tryCatch(get_opentargets_knowndrugs_detailed(g, page_size = 3000), error = function(e) {
    message("  Error: ", e$message)
    NULL
  })
})

known_drugs_subnet_RA_CD_unique_phase4 <- unique(known_drugs_subnet_RA_CD[known_drugs_subnet_RA_CD$status=="Completed" & known_drugs_subnet_RA_CD$clinical_precedence=="Approved" & !known_drugs_subnet_RA_CD$indication_is_therapeutic_area &
known_drugs_subnet_RA_CD$indication_max_phase == 4, c(1,2,5,7,12:16,24,25,26)]) %>% na.omit()
```

### Phase 4 Drug → Gene → Indication

```{r}
# known_drugs_subnet_RA_CD_unique = known_drugs_subnet_RA_CD %>% 
#   filter(is_approved, status == "Completed", drug_type %in% c("Protein", "Antibody", "Antibody drug conjugate")) %>% 
#   mutate(
#   known_drugs = case_when(is.na(drug_name) ~ "Unknown",
#                           TRUE ~ "Known drugs") 
# )

known_drugs_subnet_RA_CD_unique_approved <-
  known_drugs_subnet_RA_CD_unique_phase4 %>%
  select(
    gene,
    drug_name,
    drug_type,
    indication_name,
    indication_max_phase
  ) %>%
  mutate(
    drug_gene = paste(drug_name, gene, sep = "->"),
    value = 1
  ) %>%
  distinct() %>%
  filter(
    !str_detect(
      indication_name,
      regex("(neoplasm|cancer|carcinoma|leukemia)", ignore_case = TRUE)
    )
  )



known_drugs_subnet_RA_CD_unique_approved_mat = known_drugs_subnet_RA_CD_unique_approved %>%
  select(drug_gene, indication_name, value) %>% unique() %>% 
  pivot_wider(names_from = "drug_gene", values_from = "value", values_fill = 0) %>% 
  tibble::column_to_rownames(var = "indication_name")

rownames(known_drugs_subnet_RA_CD_unique_approved_mat) = gsub(",", "\n", rownames(known_drugs_subnet_RA_CD_unique_approved_mat))
#known_drugs_subnet_RA_CD_unique = known_drugs_subnet_RA_CD_unique[,c(1,3,5,8)]
```

```{r fig.width=42, fig.height=30}


heat_m_drug = Heatmap(t(known_drugs_subnet_RA_CD_unique_approved_mat),
        cluster_columns = T,
        name = "Indication",
        col = structure(c("grey80", "grey25"), names = c(0, 1)),
        width = nrow(known_drugs_subnet_RA_CD_unique_approved_mat) * unit(18, "mm"),
        height = ncol(known_drugs_subnet_RA_CD_unique_approved_mat) * unit(18, "mm"),
        show_row_dend = F,
        show_column_dend = F,
        rect_gp = gpar(col = "white"),
        row_dend_width = unit(3, "cm"),
        row_dend_gp = gpar(lwd =3),
        heatmap_legend_param = list(
        title_gp = gpar(fontsize = 25, fontface = "bold"),
        labels_gp = gpar(fontsize = 25),
        at = c(0, 1), 
        labels = c("NO", "Yes"),
        grid_width = unit(12, "mm"),
        grid_height = unit(12, "mm"),
        title_gap = unit(5, "mm"),
        direction = "horizontal"),
      row_names_gp = gpar(fontsize = 25),
      row_names_side = "left",
      column_names_gp = gpar(fontsize = 25),
      column_title_gp = gpar(fontsize = 30, fontface = "bold"),
      column_names_side = "top", column_title_side = "bottom",
      column_title = "Approved drugs",
      column_names_rot = 45
      )

#heat_m = ggplotify::as.ggplot(heat_m)
  
draw(heat_m_drug, heatmap_legend_side = "bottom")

```

```{r}

svg(str_glue("{figurePath}Heatmap_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_2025.svg"), width = 42, height = 30)
draw(heat_m_drug, heatmap_legend_side = "right")
dev.off()


```

#### Figure 6E: Tripartite plot

```{r fig.width=32, fig.height=20}
known_drugs_subnet_RA_CD_unique_approved_tripartite = known_drugs_subnet_RA_CD_unique_approved
known_drugs_subnet_RA_CD_unique_approved_tripartite$indication_name = str_replace(known_drugs_subnet_RA_CD_unique_approved_tripartite$indication_name, " Positive", "+") %>%
  str_to_sentence()

known_drugs_subnet_RA_CD_unique_approved_tripartite$indication_name <- case_when(str_detect(known_drugs_subnet_RA_CD_unique_approved_tripartite$indication_name, regex("hypertension", ignore_case = TRUE)) ~ "Hypertension", TRUE ~ known_drugs_subnet_RA_CD_unique_approved_tripartite$indication_name)

create_tripartite_ggplot(
  data = known_drugs_subnet_RA_CD_unique_approved_tripartite, 
  drug_col = "drug_name", 
  gene_col = "gene", 
  indication_col = "indication_name",
  weight_col = "indication_max_phase",
  color_by = "gene", gene_color = "Set1",
  node_size = 4, label_size = 7.5,
  text_size = 20,
  pathwidth = 2,
  text_face = "plain",
  trial_dot_size = 20,
  trial_dot_color = "grey50",
  title = "Approved drug → Gene → Indication Colored by Drug"
)


```

```{r}
ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_tripartite_2025.svg"), width = 32, height = 20, bg = "white")

ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_Approved_drugs_tripartite_2025.png"), width = 32, height = 20, bg = "white")
```

### Phase 3 Drug → Gene → Indication

```{r fig.width=35, fig.height=80}

known_drugs_subnet_RA_CD_unique = known_drugs_subnet_RA_CD %>% mutate(
  known_drugs = case_when(is.na(drug_name) ~ "Unknown",
                          TRUE ~ "Known drugs") 
)

known_drugs_subnet_RA_CD_unique_phase3 = known_drugs_subnet_RA_CD_unique %>% 
  select(gene, drug_name, drug_type, indication_name, indication_max_phase, status) %>% 
  filter(indication_max_phase %in% (1:3), status == "Completed") %>% 
  mutate(drug_gene = paste(drug_name, gene, sep = "->")) %>% unique() %>% mutate(value = 1)


known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3 = known_drugs_subnet_RA_CD_unique_phase3
known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3$indication_name = str_replace(known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3$indication_name, " Positive", "+") %>%
  str_to_sentence()

known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3$phase_for_indication2 = as.character(known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3$indication_max_phase)

# create_tripartite_ggplot_2(
#   known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3, 
#   drug_col = "drug_name", 
#   gene_col = "gene", 
#   indication_col = "indication_name",
#   weight_col = "indication_max_phase",
#   color_by = "phase_for_indication2",
#   color_name = "Phase for indication",
#   node_size = 4, label_size = 8.5,
#   custom_color = c("#1f77b4", "#2ca02c", "#d62728"),
#   text_size = 22,
#   pathwidth = 2,
#   text_face = "plain", line_alpha = 0.8,
#   title = "Drug → Gene → Indication - Colored by Drug (phase 1 to 3)"
# )

tripartite_ggplot3 <- create_tripartite_ggplot(
  data = known_drugs_subnet_RA_CD_unique_approved_tripartite_phase3, 
  drug_col = "drug_name", 
  gene_col = "gene", 
  indication_col = "indication_name",
  weight_col = "indication_max_phase",
  color_by = "gene", gene_color = "Set1",
  node_size = 4, label_size = 7.5,
  text_size = 20,
  pathwidth = 2,
  text_face = "plain",
  trial_dot_size = 20,
  trial_dot_color = "grey50",
  title = "Approved drug → Gene → Indication Colored by Drug"
)

tripartite_ggplot3

ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_drugs_phase_1_to_3_tripartite_2025.svg"), width = 35, height = 28, bg = "white")

ggsave(str_glue("{figurePath}Tripartite_plot_IL1B_Macr_up_Genetic_OPT_Mac_drugs_phase_1_to_3_tripartite_2025.png"), width = 35, height = 28, bg = "white")
```
### Figure 6C

```{r fig.fullwidth = TRUE, fig.width=2.5, fig.height=3.5}
drugs_ori = known_drugs_subnet_RA_CD_unique %>%
  filter(status == "Completed" | known_drugs == "Unknown") %>% 
  dplyr::select(gene, known_drugs, clinical_precedence) %>%
  mutate(known_drugs = case_when(known_drugs == "Known drugs" ~ "Yes", TRUE ~ "NO")) %>% 
  filter(gene != "CD274") %>% # remove PD-L1 (for cancer)
  unique() 
known_drugs_ori = drugs_ori %>% dplyr::group_by(known_drugs) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count)

barplot4netannot(
  data = known_drugs_ori,
  X = "known_drugs",
  Y = "Count",
  label = TRUE,
  ylab = "Number of genes",
  name = "Genes with \n known drug",
  key.size = 0.4,
  label_extra_position = 1.2,
  colmap = vega_20[c(12,11)]
)

ggsave(str_glue("{figurePath}known_drugs_subnet_RA_CD_J_unique_2025.png"), bg='white')
```

### Figure 6D


```{r fig.fullwidth = TRUE, fig.width=3, fig.height=3.5}
# known_drugs_ori_2 = known_drugs_subnet_RA_CD_unique[known_drugs_subnet_RA_CD_unique$known_drugs == "Known drugs",] %>% 
#   filter(status == "Completed") %>% 
#   dplyr::select(gene, clinical_precedence, status, indication_name) %>% 
#   na.omit() %>% unique()# %>% 
known_drugs_ori_2 = drugs_ori %>% na.omit() %>% 
  filter(known_drugs == "Yes") %>% 
  dplyr::group_by(clinical_precedence) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count) %>% 
  dplyr::mutate(clinical_precedence = as.factor(clinical_precedence))

known_drugs_ori_2$clinical_precedence = factor(known_drugs_ori_2$clinical_precedence,
                                               levels = c("Phase I", "Phase II", "Phase III", "Approved"))

barplot4netannot(
  data = known_drugs_ori_2,
  X = "clinical_precedence",
  Y = "Count",
  label = TRUE,
  name = "Clinical Trial \n Precedence",
  key.size = 0.4,
  label_extra_position = 0.3,
  colmap =  c("#e7e7e7", "#D4B996", "#99847A", "#3A405A")#vega_20
)

ggsave(str_glue("{figurePath}known_drugs_subnet_RA_CD_J_Phase_2025.png"), bg='transparent')
```

## Prioritize shared module

```{r}

subnet_RA_CD_priori = igraph::subgraph(graph = net, vids = (V(net)$Origin != "Intermediate"))

#degree=igraph::degree(subnet_RA_CD_priori)

#subnet_RA_CD_priori = set.vertex.attribute(graph=subnet_RA_CD_priori, name="degree", value=as.numeric(degree))

#V(subnet_RA_CD_priori)$degree = degree

#subnet_RA_CD_priori = igraph::subgraph(graph = subnet_RA_CD_priori, vids = (degree != 0))

page.rank.net.priori = page_rank(
  subnet_RA_CD_priori,
  personalized = as.numeric(V(subnet_RA_CD_priori)$Score)
)


subnet_RA_CD_priori = set_vertex_attr(graph=subnet_RA_CD_priori, name="page.rank", value=page.rank.net.priori$vector)


cwt.net.priori=cluster_walktrap(subnet_RA_CD_priori, 
                           weights = E(subnet_RA_CD_priori)$weight, 
                           steps = 5,
                           merges = F,
                           modularity = TRUE, 
                           membership = TRUE)

table(cwt.net.priori$membership)
  

subnet_RA_CD_priori = set_vertex_attr(graph=subnet_RA_CD_priori, name="cwt", value=as.factor(cwt.net.priori$membership))

#degree=igraph::degree(subnet_RA_CD_priori)

#subnet_RA_CD_priori = set_vertex_attr(graph=subnet_RA_CD_priori, name="degree", value=degree)


#subnet_RA_CD_meta.priori = igraph::as_data_frame(subnet_RA_CD_priori, what = "vertices")

class(V(subnet_RA_CD_priori))


```

```{r fig.fullwidth = TRUE, fig.width=8, fig.height=6}
ggraph(subnet_RA_CD_priori, layout = "kk") +
  geom_edge_link0(width=0.2, colour = "gray90") +
  #scale_color_brewer(palette = "Paired") +
  geom_edge_fan(aes(alpha = after_stat(index)), show.legend = FALSE) +
  geom_node_point(aes(color = cwt, size = page.rank)) +
  geom_node_text(aes(label = name), repel = T) +
  theme_graph()
```

### Known drug targets

```{r}

known_drugs_subnet_RA_CD_priori = gene2knowndrugs(geneid = as_ids(V(subnet_RA_CD_priori)), ensId = FALSE)

known_drugs_subnet_RA_CD_priori = unique(known_drugs_subnet_RA_CD_priori)
```

```{r}

known_drugs_subnet_RA_CD_unique_priori = known_drugs_subnet_RA_CD_priori %>% mutate(known_drugs = case_when(is.na(drug_id) ~ "Unknown",                           TRUE ~ "Known drugs")  )  

known_drugs_subnet_RA_CD_unique_priori = known_drugs_subnet_RA_CD_unique_priori[,c(1,3,5,8)]
```

```{r fig.fullwidth = TRUE, fig.width=2.5, fig.height=3.5}

known_drugs_ori_priori = known_drugs_subnet_RA_CD_unique_priori %>% dplyr::select(gene.name, known_drugs) %>%
  mutate(known_drugs = case_when(known_drugs == "Known drugs" ~ "Yes", TRUE ~ "NO")) %>%
  unique() %>% dplyr::group_by(known_drugs) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count)  

barplot4netannot(data = known_drugs_ori_priori, X = "known_drugs", Y = "Count", label = TRUE, ylab = "Number of genes", name = "Drugged genes", key.size = 0.4,   colmap = vega_20[11:12])

#ggsave(str_glue("{figurePath}known_drugs_subnet_RA_CD_J_unique_prioritized.png"), bg='transparent')
```

```{r fig.fullwidth = TRUE, fig.width=3, fig.height=3.5}


known_drugs_ori_2_priori = known_drugs_subnet_RA_CD_unique_priori[known_drugs_subnet_RA_CD_unique_priori$known_drugs == "Known drugs",] %>%
  unique() %>% dplyr::select(ClinicalTrialPhase) %>% na.omit() %>%
  dplyr::group_by(ClinicalTrialPhase) %>% 
  dplyr::summarise(Count = n()) %>% dplyr::arrange(Count) %>% 
  dplyr::mutate(ClinicalTrialPhase = as.factor(ClinicalTrialPhase))

barplot4netannot(data = known_drugs_ori_2_priori, X = "ClinicalTrialPhase", Y = "Count", ylab = "Number of clinical trials", label = TRUE, name = "Clinical Trial\nPhase", key.size = 0.4, colmap = vega_20 )  

#ggsave(str_glue("{figurePath}known_drugs_subnet_RA_CD_J_Phase_prioritized.png"), bg='transparent')
```

## Known drug targets all gene

```{r}
#library(biomaRt)
#known_drugs_subnet_RA_CD = gene2knowndrugs(geneid = as_ids(V(net)), ensId = FALSE)


known_drugs_subnet_RA_CD_all <- map_df(rownames(IL1B_macr_up_wide), function(g) {
  Sys.sleep(0.5)
  tryCatch(get_opentargets_knowndrugs(g, page_size = 3000), error = function(e) {
    message("  Error: ", e$message)
    NULL
  })
})
```

## GSEA

```{python}
import os,sys
import numpy as np
import pandas as pd
#import drug2cell as d2c
#import scanpy as sc
import sctk
#import gc
#sys.path.append("/mnt/c/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

sys.path.append("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

from inflapy.immunoFunc import *
import gseapy as gp
```

```{python}
gp.get_library_name()
```

```{r}
genelist = as_ids(V(net))
```

```{python}
#gene_sets=['MSigDB_Hallmark_2020', 'Reactome_2022']
genelist = r.genelist
gene_sets=['MSigDB_Hallmark_2020']

enr = gp.enrichr(gene_list=genelist,
                     gene_sets=gene_sets,
                     organism='human', 
                     outdir=None, # don't write to disk
                    )
enr1 = enr.results
enr1

```

```{r}
enr1 = py$enr1

```

```{python}
#gene_sets=['MSigDB_Hallmark_2020', 'Reactome_2022']
genelist = r.genelist
gene_sets=['Reactome_2022']

enr = gp.enrichr(gene_list=genelist,
                     gene_sets=gene_sets,
                     organism='human', 
                     outdir=None, # don't write to disk
                    )
enr2 = enr.results
enr2
```

```{r}
enr2 = py$enr2

```

```{r  fig.fullwidth = TRUE, fig.width=6, fig.height=2.5}
gsea_plot(gsea=enr1, n_top = 10, x = "Term", y = "Adjusted P-value", log_y = F, bar_plot = TRUE, ylab="Adjusted P-value", xlab = "Hallmark GSEA")

#ggsave(str_glue("{figurePath}GSEA_Bar_Tissue_IntermediateMacr_RA_CD_subgraph_Modules_Jaccard_index_Hallmark_2020.png"), bg='transparent', width=6, height=2.5)
```

```{r}
enr2$Term = str_remove(enr2$Term, " R-HSA-.*$") %>% str_wrap(., 35)
```

```{r  fig.fullwidth = TRUE, fig.width=8, fig.height=4}
gsea_plot(
  gsea = enr2,
  n_top = 10,
  x = "Term",
  y = "Adjusted P-value",
  log_y = T,
  bar_plot = TRUE,
  ylab = "log2(Adjusted P-value)",
  xlab = "Reactome")


#ggsave(str_glue("{figurePath}GSEA_Bar_Tissue_IntermediateMacr_RA_CD_subgraph_Modules_Jaccard_index_Reactome.png"), bg='transparent',
       #width=8, height=3.5)
```

## Save

```{r}
save.image("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/All_Tissue_or_Nonimmune/Subset_Reintegrated_full/Milo/individual_disease_nhoo/Tissue_Myeloid_NetExpansion_from_genetic_OPT_session_2025.RData")
```

```{r}
#load("G:/My Drive/Projects/SCVI_Results/All_Tissue_or_Nonimmune/Subset_Reintegrated_full/Milo/individual_disease_nhoo/Tissue_Myeloid_NetExpansion_from_genetic_OPT_session.RData")

load("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/All_Tissue_or_Nonimmune/Subset_Reintegrated_full/Milo/individual_disease_nhoo/Tissue_Myeloid_NetExpansion_from_genetic_OPT_session_2025.RData")
```
