---
title: "Differential expression of Milo nhoods"
author: "Bidossessi Hounkpe"
date: "04/08/2023"
format: html
  df-print: kable
editor: visual
---

```{r}
rm(list = ls())
```

```{r}
library(tidyverse)
library(SingleCellExperiment)
library(Matrix)
library(scran)
library(scuttle)
library(glmGamPoi)
library(edgeR)
library(ggbeeswarm)
library(ComplexHeatmap)
library(reticulate)
library(ggrepel)

basepath <- "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/scrnaseq-autoimmune-atlas/notebooks/4-R/"

# DEG

source(str_glue(basepath, "deg_fun.R"))

#source("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/Integrated/R/colors.R")
source(str_glue(basepath, "colors.R"))
# Source helper functions
#source("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/Integrated/R/helper_fun.R")
source(str_glue(basepath, "helper_fun.R"))
# Network expansion 
#source("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/Integrated/R/openTarget.R")
source(str_glue(basepath, "openTarget.R"))

source(str_glue(basepath, "Milo_horizontal_plot.R"))
# Network expansion 

#source("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/Integrated/R/openTarget.R")
figurePath = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Figures/"


degPathEnriched = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/DEG/Enriched/"

degPathDepleted = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/DEG/Depleted/"
```

```{r}
#Use conda env for reticulate
reticulate::use_condaenv("ScGwas", required = TRUE)
```

```{python}
import os,sys
import numpy as np
import pandas as pd
#import drug2cell as d2c
#import scanpy as sc
#import sctk
#import gc
#sys.path.append("/mnt/c/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

sys.path.append("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/scripts/autoimmuneSC/AIID_notebooks/")

#from inflapy.immunoFunc import *
import gseapy as gp
```

```{python}
print(sys.executable)
"/Users/bidos/miniconda3/envs/ScGwas/pyt"
#import blitzgsea as blitz
import platform
print(platform.python_version())
import blitzgsea as blitz
```

```{python}

# read signature as pandas dataframe
signature = pd.read_csv("https://github.com/MaayanLab/blitzgsea/raw/main/testing/ageing_muscle_gtex.tsv")

# list available gene set libraries in Enrichr
blitz.enrichr.print_libraries()

# use enrichr submodule to retrieve gene set library
library = blitz.enrichr.get_library("KEGG_2021_Human")

# run enrichment analysis
result = blitz.gsea(signature, library)

```

```{r}
reticulate::py_install("blitzgsea")
```

# Umap Myeloid from tissue

```{r}
obs = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/Myeloid_From_PBMC_label_transfer_Milo_3K.csv")

umap = read.csv("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/Myeloid_From_PBMC_label_transfer_Milo_3K_UMAP.csv")
```

```{r}
unique(umap$Cell_type_scANVI)
```
### Figure 3A

```{r fig.width= 10, fig.height=7}


umap = umap[umap$Cell_type_scANVI != "CCR7_CDE3",]

umap$Cell_type_scANVI = case_when(
  umap$Cell_type_scANVI == "cMon_1" ~ "Activated cMon",
  umap$Cell_type_scANVI == "cMon_2" ~ "cMon",
  TRUE ~ umap$Cell_type_scANVI
)

umap %>% filter(Cell_type_scANVI != "Unknown") %>% 
  ggplot(aes(x=umap_1, y=umap_2, color = Cell_type_scANVI)) +
  geom_point() +
  theme_bw() +
  scale_colour_manual(values = color_cell_pbmc)
  
```

```{r fig.width= 5, fig.height=4}
plot_umap(
  umap_data = umap,
  cluster = "Cell_type_scANVI",
  umap_1 = "umap_1",
  umap_2 = "umap_2",
  selected_clusters = c(
    "Activated cMon",
    "cMon",
    "Non-cMon"
  ),
  shape.size = 0.5,
  cluster_palette = color_cell_pbmc,
  arrow.length = 3,
  repel_text_size = 3
)



ggsave(
  str_glue(
    "{figurePath}Umap_Milo_myel_From_PBMC_Only_cMon_Non_cMon.svg"
  ),
  width = 5,
  height = 5,
  bg = 'transparent'
)


ggsave(
  str_glue(
    "{figurePath}Umap_Milo_myel_From_PBMC_Only_cMon_Non_cMon.png"
  ),
  width = 5,
  height = 4,
  bg = 'transparent'
)
  
```


### Figure 3C: Milo plot

```{r fig.fullwidth = TRUE, fig.width=20, fig.height=10.5}

h5ad_path = 'C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/'

compare_disease_df = read.csv(str_glue("{h5ad_path}Myeloid_From_PBMC_label_transfer_Milo_3K.csv"))

#compare_disease_df = compare_disease_df[!compare_disease_df$disease %in% c("all_CD_and_spondoarthritis", "all_spondoarthritis"), ]

unique(compare_disease_df$disease)
compare_disease_df$Sig = ifelse(compare_disease_df$Sig == "True", TRUE, FALSE)

print(table(compare_disease_df$disease))

#### Prepare data


nhoo_prep = milo_prep_data(nhood_res=compare_disease_df, condition = "disease", Sig = "Sig", thres = 0.1,
                     dis_pref = "all_", index_cell = "index_cell", order_by_kth = FALSE,
                     nhood_annotation="nhood_annotation", n_shared=0,
                     SpatialFDR = "SpatialFDR", logFC = "logFC")


nhoo_prep$condition = str_replace(nhoo_prep$condition, "CD", "IBD") %>% as.factor()

nhoo_prep$condition = str_replace(nhoo_prep$condition, "hashimoto", "Hashimoto's thyroiditis")

nhoo_prep$condition = str_replace(nhoo_prep$condition, "spondoarthritis", "Spondlyoarthritis")

nhoo_prep$condition = str_replace(nhoo_prep$condition, "sjo_syndr", "Sjögren's syndrome")

nhoo_prep$nhood_annotation = case_when(
  nhoo_prep$nhood_annotation == "cMon_1" ~ "Activated cMon",
  nhoo_prep$nhood_annotation == "cMon_2" ~ "cMon",
  TRUE ~ nhoo_prep$nhood_annotation
)

milo_plot_horizontal(
  nhoo_prep = nhoo_prep[nhoo_prep$nhood_annotation != "CCR7_CDE3", ],
  rel_widths = c(0.15, 1, 0.25),
  anot_bar_height = 0.05,
  col.pal = color_cell_pbmc,
  label_size = 30,
  annot_size = 15,
  annot_col = "black",
  annot.bar.height = 0.5,
  expand_y_sc = 0.1,
  #dendro.width = 3,
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(size = 40),
  legend.title = element_text(size = 40, face = "bold"),
  legend.position = "bottom",
  legend.box.spacing = unit(5, 'cm'),
  legend.spacing =5,
  fc_legend.size =5,
  fc_annot_size = 40,
  legend.margin.custom = -180,
  dendro.width = 2, bar.title = "Blood",
  y.axis.label = "Comparative neighbourhood \nabundance",
  y.axis.label.size = 35
  
) +
  guides(fill = guide_legend(title.position="top", title.hjust = 0.5, ncol = 2))

#milo_hori_plot


```

```{r}
ggsave(str_glue("{figurePath}Milo_myel_From_PBMC_compareDisease_with_specific_spondyloarthtitis_2026.png"), width = 20, height = 10.5, bg='transparent')
ggsave(str_glue("{figurePath}Milo_myel_From_PBMC_compareDisease_with_specific_spondyloarthtitis_2026.svg"), width = 20, height = 10.5, bg='transparent')
```

```{r}
# ggsave(str_glue("{figurePath}Milo_myel_From_PBMC_compareDisease_with_specific_healthy.png"), width = 20, height = 8, bg='transparent')
# ggsave(str_glue("{figurePath}Milo_myel_From_PBMC_compareDisease_with_specific_healthy.svg"), width = 20, height = 8, bg='transparent')
```


### Data loading and preparation

```{r}
#sce_path = "/nfs/team205/bh14/Datasets/Remapped/raw_adata/SCVI_Results/All_PBMC_or_Nonimmune/Subset_Reintegrated_full/Milo/enriched_depleted_SCE/"

sce_path = "C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/"

# 

load(str_glue("{sce_path}Pseudo_test_adata_Myeloid_PBMC.RData"))

test_adata_$Cell_type_scANVI = as.character(test_adata_$Cell_type_scANVI)

test_adata_ = test_adata_[, test_adata_$Cell_type_scANVI != "CCR7_CDE3"]

test_adata_$Cell_type_scANVI = gsub(" ", "_", test_adata_$Cell_type_scANVI) %>% factor()

test_adata_$Cell_type_scANVI = gsub("Mon_SIRPA", "cMon_SIRPA", test_adata_$Cell_type_scANVI) %>% factor()

test_adata_$donor_id = factor(gsub(" ", "_", test_adata_$donor_id))

test_adata_$donor_id = factor(gsub("-", "_", test_adata_$donor_id))

test_adata_$disease_short = droplevels(test_adata_$disease_short)

unique(test_adata_$Cell_type_scANVI)
```

```{r}
table(test_adata_$disease_short, test_adata_$Cell_type_scANVI)
```

```{r}
table(test_adata_$nhoo_test_condition, test_adata_$Cell_type_scANVI)
```
## End
```{r}
my_obs = read.csv("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/PBMC_Myeloid_individual_disease_Milo_adata_reajusted.csv")

table(my_obs$Cell_type_scANVI)



```

```{r}
my_obs$disease_short = str_remove(my_obs$disease_short, "all_")
```

```{r}
my_obs$disease_dataset=paste0(my_obs$disease_short, "_", my_obs$dataset_id)
```

```{r}
table(my_obs$disease_dataset, my_obs$Cell_type_scANVI)
```

#### Abundance Plot

```{r}
levels(test_adata_$Cell_type_scANVI)
```

```{r}
table(my_obs$dataset_id, my_obs$disease_short)
```

```{r}
my_obs$disease_short = str_replace(my_obs$disease_short, "CD", "IBD") %>% as.factor()
my_obs$disease_short = str_replace(my_obs$disease_short, "hashimoto", "Hashimoto's thyroiditis")

my_obs$disease_short = str_replace(my_obs$disease_short, "spondoarthritis", "Spondyloarthritis")

my_obs$disease_short = str_replace(my_obs$disease_short, "sjo_syndr", "Sjögren's syndrome")
unique(my_obs$disease_short)
```

```{r}
my_obs$Cell_type_scANVI = case_when(
  my_obs$Cell_type_scANVI == "cMon_1" ~ "Activated cMon",
  my_obs$Cell_type_scANVI == "cMon_2" ~ "cMon",
  TRUE ~ my_obs$Cell_type_scANVI
)


```

```{r fig.width=8}
my_obs[my_obs$dataset_id %in% c("GSE157278_Hong_2020", "GSE163314_Lefferts_2021", "Perez_2022") & 
         my_obs$Cell_type_scANVI != "CCR7_CDE3",] %>% 
dplyr::select(dataset_id, disease_short, Cell_type_scANVI) %>% mutate(dataset_id = str_remove(dataset_id, "GSE[0-9]*_{1}")) %>% 
  group_by(dataset_id, Cell_type_scANVI, disease_short) %>% mutate(count = n()) %>% unique() %>%
  ggplot(aes(fill=Cell_type_scANVI, y=count, x= reorder(disease_short, count))) + 
  geom_bar(position="stack", stat="identity", color = "white") + 
  labs(x = "", y = "Composition", fill = "Cellular compartment") +
  facet_wrap(~ dataset_id, scales = "free", ncol = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_bw() + #coord_flip() +
  scale_fill_manual(values = color_cell_pbmc) +
  scale_x_reordered() +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 0.5,
      size = 12,
      angle = 90
    ),
    axis.text = element_text(size = 18, color = "black"),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 18),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.8, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.8, "cm"),
    plot.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    legend.title = element_text(size = 18)
  )


ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC_across_studies_with_healthy.svg"), width = 8, bg='transparent')

ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC_across_studies_with_healthy.png"), width = 8, bg='transparent')

```

```{r fig.fullwidth = TRUE, fig.width=9, fig.height=9}
my_obs_0_2 = my_obs[my_obs$dataset_id == "Perez_2022" & my_obs$Cell_type_scANVI != "CCR7_CDE3",]


my_obs_0_2$disease_short = as.factor(my_obs_0_2$disease_short)

my_obs_0_2$donor_nhoo = paste0(my_obs_0_2$donor_id, "_", my_obs_0_2$nhoo_test_condition)

my_obs_0_2 = my_obs_0_2 %>% dplyr::select(donor_id, donor_nhoo, disease_short, Cell_type_scANVI, nhoo_test_condition) %>% 
  # Number of cells per donor in each cell type
  group_by(donor_nhoo, Cell_type_scANVI, disease_short) %>% mutate(total_per_donor_per_celltype = n()) %>% unique() %>%  
  # total number of cells per donor
  group_by(donor_id) %>% mutate(Cell_count_per_donor=sum(total_per_donor_per_celltype)) %>% unique() %>% ungroup() %>% 
  # Cells per 1000
  mutate(Cell_per_1000=total_per_donor_per_celltype*1000/Cell_count_per_donor) %>% unique() %>% 
  group_by(donor_id) %>% mutate(dsum = sum(Cell_per_1000)) %>% 
  # number of cell per nhood per disease
  group_by(disease_short, Cell_type_scANVI, nhoo_test_condition) %>% mutate(norm_Cell_count_per_nhood_disease=sum(Cell_per_1000))

#some donors have only one cell type (Cell_per_1000==1000)

my_obs_0_3 = unique(my_obs_0_2[,c(3:5,10)])

ggplot(my_obs_0_3, aes(fill=nhoo_test_condition, y=norm_Cell_count_per_nhood_disease, x=reorder_within(x = disease_short, by = norm_Cell_count_per_nhood_disease, within = Cell_type_scANVI, fun = sum))) + 
  geom_bar(position="stack", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Abundance", fill = "Cellular \n compartment") +
    #scale_y_continuous(expand = c(0,1)) +
    #scale_fill_manual(values=celcolormap) +
  facet_wrap(~ Cell_type_scANVI, scales = "free", ncol = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_bw() + #coord_flip() +
  scale_x_reordered() +
  #geom_text(aes(label = after_stat(y), group = disease_short), 
            #stat = 'summary', fun = sum, vjust = -0.2, hjust = 0.5) +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 0.5,
      size = 12,
      angle = 90
    ),
    axis.text = element_text(size = 18, color = "black"),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 18),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.8, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.8, "cm"),
    plot.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    legend.title = element_text(size = 18)
  )


ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PMBC_compareDisease_SLE.svg"), width = 9, height = 9, bg='transparent')

ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PBMC_compareDisease_SLE.png"), width = 9, height = 9, bg='transparent')

```

```{r fig.fullwidth = TRUE, fig.width=12, fig.height=10}
my_obs_0_2 = my_obs[my_obs$dataset_id != "Perez_2022" & my_obs$Cell_type_scANVI != "CCR7_CDE3",]


my_obs_0_2$disease_short = as.factor(my_obs_0_2$disease_short)

my_obs_0_2$donor_nhoo = paste0(my_obs_0_2$donor_id, "_", my_obs_0_2$nhoo_test_condition)

my_obs_0_2 = my_obs_0_2 %>% dplyr::select(donor_id, donor_nhoo, disease_short, Cell_type_scANVI, nhoo_test_condition) %>% 
  # Number of cells per donor in each cell type
  group_by(donor_nhoo, Cell_type_scANVI, disease_short) %>% mutate(total_per_donor_per_celltype = n()) %>% unique() %>%  
  # total number of cells per donor
  group_by(donor_id) %>% mutate(Cell_count_per_donor=sum(total_per_donor_per_celltype)) %>% unique() %>% ungroup() %>% 
  # Cells per 1000
  mutate(Cell_per_1000=total_per_donor_per_celltype*1000/Cell_count_per_donor) %>% unique() %>% 
  group_by(donor_id) %>% mutate(dsum = sum(Cell_per_1000)) %>% 
  # number of cell per nhood per disease
  group_by(disease_short, Cell_type_scANVI, nhoo_test_condition) %>% mutate(norm_Cell_count_per_nhood_disease=sum(Cell_per_1000))

#some donors have only one cell type (Cell_per_1000==1000)

my_obs_0_3 = unique(my_obs_0_2[,c(3:5,10)])

ggplot(my_obs_0_3, aes(fill=nhoo_test_condition, y=norm_Cell_count_per_nhood_disease, x=reorder_within(x = disease_short, by = norm_Cell_count_per_nhood_disease, within = Cell_type_scANVI, fun = sum))) + 
  geom_bar(position="stack", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Abundance", fill = "Cellular \n compartment") +
    #scale_y_continuous(expand = c(0,1)) +
    #scale_fill_manual(values=celcolormap) +
  facet_wrap(~ Cell_type_scANVI, scales = "free", ncol = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_bw() + #coord_flip() +
  scale_x_reordered() +
  #geom_text(aes(label = after_stat(y), group = disease_short), 
            #stat = 'summary', fun = sum, vjust = -0.2, hjust = 0.5) +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 0.5,
      size = 12,
      angle = 90
    ),
    axis.text = element_text(size = 18, color = "black"),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 18),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.8, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.8, "cm"),
    plot.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    legend.title = element_text(size = 18)
  )


ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PMBC_compareDisease.svg"), width = 12, height = 10, bg='transparent')

ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PBMC_compareDisease.png"), width = 12, height = 10, bg='transparent')

```

```{r fig.fullwidth = TRUE, fig.width=18, fig.height=16}
my_obs_0_2 = my_obs[my_obs$Cell_type_scANVI != "CCR7_CDE3",]


my_obs_0_2$disease_short = as.factor(my_obs_0_2$disease_short)

my_obs_0_2 = my_obs_0_2 %>% dplyr::select(donor_id, disease_short, Cell_type_scANVI) %>% 
  # Number of cells per donor in each cell type
  group_by(donor_id, Cell_type_scANVI, disease_short) %>% mutate(total_per_donor_per_celltype = n()) %>% unique() %>%  
  # total number of cells per donor
  group_by(donor_id) %>% mutate(Cell_count_per_donor=sum(total_per_donor_per_celltype)) %>% unique() %>% ungroup() %>% 
  # Cells per 1000
  mutate(Cell_per_1000=total_per_donor_per_celltype*1000/Cell_count_per_donor) %>% unique()

#some donors have only one cell type (Cell_per_1000==1000)

ggplot(my_obs_0_2, aes(y=Cell_per_1000/10, x=reorder_within(x = disease_short, by = Cell_per_1000, within = Cell_type_scANVI, fun = mean))) + 
    geom_boxplot() +
    geom_beeswarm(stat="identity") + 

    labs(x = "Tissue", y = "Percentage of cell per donor", fill = "DA") +
    facet_wrap(~ Cell_type_scANVI, scales = "free", ncol = 4) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    theme_bw() + #coord_flip() + 
    scale_x_reordered() +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 1,
      size = 10,
      angle = 45
    ),
    
    axis.text = element_text(size = 18, color = "black"),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 18),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.8, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.8, "cm"),
    plot.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    legend.title = element_text(size = 18)
  )


ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PBMC_compareDisease.svg"), width = 18, height = 16, bg='transparent')

ggsave(str_glue("{figurePath}Cell_per_donor_Milo_myel_From_PBMC_compareDisease.png"), width = 18, height = 16, bg='transparent')

```

##### Figure 3B: Cell abundance across diseases

```{r fig.fullwidth = TRUE, fig.width=12, fig.height=4}
#myobs = my_obs %>% filter(disease_short!="Healthy") %>% group_by(Cell_type_scANVI, disease_short) %>% summarise(Cell_count=n())
myobs_1 = my_obs[my_obs$Cell_type_scANVI != "CCR7_CDE3",] %>% group_by(Cell_type_scANVI, disease_short) %>% summarise(Cell_count=n())

#my_obs$tissue = ifelse(my_obs$tissue == "Macrophage", "Synovium", as.character(my_obs$tissue))


myobs_1$disease_short = factor(myobs_1$disease_short, levels = c(sort(as.character(
  unique(myobs_1$disease_short[myobs_1$disease_short != "Healthy"])
), decreasing = T), "Healthy"))
#head(my_obs)
celcompodisease = ggplot(myobs_1, aes(fill=Cell_type_scANVI, y=Cell_count, x=disease_short)) + 
    geom_bar(position="fill", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Composition (%)", fill = "Cell compartment") +
    scale_y_continuous(expand = c(0,0)) +
    scale_fill_manual(values=color_cell_pbmc) +
    theme_classic() + coord_flip() + 
    theme(axis.text.x = element_text(hjust=1, vjust = 0.5, size = 22, angle = 90),
          axis.text.y = element_text(size=22),
          axis.title = element_text(size=22),
         legend.key.size = unit(1, 'cm'),
         legend.text = element_text(size=22),
         legend.title = element_text(size=22),
         axis.line = element_line(colour = "black", linewidth = 0.6),
         legend.key.width = unit(0.6, "cm"),
         legend.key.height = unit(0.6, "cm")) #+ scale_colour_brewer(palette = "Paired") 
    
celcompodisease


ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC.svg"), width = 10, height = 4, bg='transparent')

ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC.png"), width = 12, height = 4, bg='transparent')
```

### Figure 3BX´

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=6}
#myobs = my_obs %>% filter(disease_short!="Healthy") %>% group_by(Cell_type_scANVI, disease_short) %>% summarise(Cell_count=n())
myobs_1 = my_obs[my_obs$Cell_type_scANVI != "CCR7_CDE3",] %>% group_by(Cell_type_scANVI, disease_short) %>% summarise(Cell_count=n())

#my_obs$tissue = ifelse(my_obs$tissue == "Macrophage", "Synovium", as.character(my_obs$tissue))

myobs_1$disease_short = factor(myobs_1$disease_short, levels = c("Healthy", sort(as.character(
  unique(myobs_1$disease_short[myobs_1$disease_short != "Healthy"])
))))
#head(my_obs)
celcompodisease = ggplot(myobs_1, aes(fill=Cell_type_scANVI, y=Cell_count, x=disease_short)) + 
    geom_bar(position="fill", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Composition (%)", fill = "Cell compartment") +
    scale_y_continuous(expand = c(0,0)) +
    scale_fill_manual(values=color_cell_pbmc) +
    theme_classic() +# coord_flip() + 
    theme(axis.text.x = element_text(hjust=1, vjust = 0.5, size = 18, angle = 90),
          axis.text.y = element_text(size=18),
          axis.title = element_text(size=18),
         legend.key.size = unit(1, 'cm'),
         legend.text = element_text(size=18),
         legend.title = element_text(size=18),
         axis.line = element_line(colour = "black", linewidth = 0.6),
         legend.key.width = unit(0.6, "cm"),
         legend.key.height = unit(0.6, "cm")) #+ scale_colour_brewer(palette = "Paired") 
    
celcompodisease


#ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC_vertical.svg"), width = 5, height = 5, bg='transparent')

ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC_vertical.png"), width = 5, height = 6, bg='transparent')
```

```{r fig.fullwidth = TRUE, fig.width=7, fig.height=6}
myobs = my_obs %>% filter(disease_short!="Healthy") %>% group_by(Cell_type_scANVI, disease_short) %>% summarise(Cell_count=n())

celcompodisease2 = ggplot(myobs, aes(fill=Cell_type_scANVI, y=Cell_count, x=reorder(disease_short, Cell_count))) + 
    geom_bar(position="stack", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Composition", fill = "Cellular \n compartment") +
    #scale_y_log10(expand = c(0,1)) +
    scale_fill_manual(values=color_cell_pbmc) +
    theme_bw() + #coord_flip() +
    geom_text(aes(label = after_stat(y), group = disease_short), 
              stat = 'summary', fun = sum, vjust = -0.2, hjust = 0.5) +
    theme(axis.text.x = element_text(hjust=1, vjust = 0.5, size = 12, angle = 90),
          axis.text.y = element_text(size=12, color = "black"),
         legend.key.size = unit(0.6, 'cm'),
         legend.text = element_text(size=10),
         #axis.line = element_line(colour = "black", linewidth = 0.6),
         legend.key.width = unit(0.4, "cm"),
         axis.ticks.x = element_blank(),
         panel.grid = element_blank(),
         legend.key.height = unit(0.4, "cm")) #+ scale_colour_brewer(palette = "Paired") 
    
celcompodisease2


ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Disease_PBMC_Stack.svg"), bg='transparent')
```

### Healthy tissue

```{r fig.fullwidth = TRUE, fig.width=4, fig.height=6}

myobs_ctrl = my_obs %>% filter(disease_short=="Healthy") %>% group_by(Cell_type_scANVI, dataset_id) %>% summarise(Cell_count=n())


celcompocontrol = ggplot(myobs_ctrl, aes(fill=Cell_type_scANVI, y=Cell_count, x=reorder(dataset_id, Cell_count))) + 
    geom_bar(position="stack", stat="identity") + 
    #geom_col(width = 0.4) +
    labs(x = "", y = "Composition", fill = "Cellular \n compartment") +
    scale_y_log10(expand = c(0,0)) +
    scale_fill_manual(values=color_cell_pbmc) +
    theme_bw() + #coord_flip() +
    #geom_text(aes(label = after_stat(y), group = PBMC_harmo), 
             # stat = 'summary', fun = sum, vjust = -0.2, hjust = 0.5) +
    theme(axis.text.x = element_text(hjust=1, vjust = 0.5, size = 12, angle = 90),
          axis.text.y = element_text(size=12, color = "black"),
         legend.key.size = unit(0.6, 'cm'),
         legend.text = element_text(size=10),
         #axis.line = element_line(colour = "black", linewidth = 0.6),
         legend.key.width = unit(0.4, "cm"),
         axis.ticks.x = element_blank(),
         panel.grid = element_blank(),
         legend.key.height = unit(0.4, "cm")) #+ scale_colour_brewer(palette = "Paired") 
    
celcompocontrol


ggsave(str_glue("{figurePath}/Myeloid_cell_abundance_Healthy_control_PBMC_Macrophage_Stack.svg"), bg='transparent')
```

```{r}
unique(my_obs$Cell_type_scANVI)
```

```{r}
unique(my_obs$disease_short)
```

##### DC1

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=4}


plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="DC1",], disease_column = "disease_short", legend_title="Disease", xlab = "", ylab = "DC1 Cells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_DC1.svg"), width = 5, height = 4, bg='transparent')
```

##### DC2

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=4}

# "Activated cMon"    "CCR7_CDE3" "cMon"    "DC2"       "Non-cMon"  "Mon_SIRPA" "DC1"       "pDC"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="DC2",], disease_column = "disease_short", legend_title="Disease",
                             xlab = "", ylab = "DC2 Cells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_DC2.svg"), width = 5, height = 4, bg='transparent')
```

##### Non cMon

```{r fig.fullwidth = TRUE, fig.width=6, fig.height=4}


# "Activated cMon"    "CCR7_CDE3" "cMon"    "DC2"       "Non-cMon"  "Mon_SIRPA" "DC1"       "pDC"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="Non-cMon",], disease_column = "disease_short", legend_title="Disease", xlab = "", ylab = "Non-cMon \nCells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_Non_cMon.svg"), width = 6, height = 4, bg='transparent')
```

##### pDC

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=4}


# "Activated cMon"    "CCR7_CDE3" "cMon"    "Mon_SIRPA" "DC1"       "pDC"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="pDC",], disease_column = "disease_short", legend_title="Disease",
                             xlab = "", ylab = "pDC Cells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_pDC.svg"), width = 5, height = 4, bg='transparent')
```

##### Activated cMon

```{r fig.width=8, fig.height=3}
#[my_obs$dataset_id %in% c("GSE157278_Hong_2020", "GSE163314_Lefferts_2021", "Perez_2022") &

my_obs$Cell_type_scANVI = as.factor(my_obs$Cell_type_scANVI)

my_obs[my_obs$Cell_type_scANVI=="Activated cMon",] %>% 
dplyr::select(nhoo_test_condition, dataset_id, disease_short, Cell_type_scANVI) %>% mutate(dataset_id = str_remove(dataset_id, "GSE[0-9]*_{1}"),
                                                                                     disease_short = str_remove(disease_short, "all_")) %>% 
  group_by(nhoo_test_condition, dataset_id, Cell_type_scANVI, disease_short) %>% mutate(count = n()) %>% unique() %>%
  ggplot(aes(fill=nhoo_test_condition, y=count, x= reorder(disease_short, count))) + 
  geom_bar(position="stack", stat="identity", color = "white") + 
  labs(x = "", y = "Nhood Composition", fill = "nhood") +
  facet_wrap(~ dataset_id, scales = "free", ncol = 5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_bw() + #coord_flip() +
  scale_fill_manual(values = c(Enriched = "#459395", Depleted = "#FDA638", Non_sig = "grey50")) + #459395 EB7C69 FDA638
  scale_x_reordered() +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 0.5,
      size = 12,
      angle = 90
    ),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.text.x.bottom = element_blank(),
    legend.key.size = unit(0.6, 'cm'),
    legend.text = element_text(size = 10),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.4, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.4, "cm")
  )

ggsave(str_glue("{figurePath}Nhood_abundance_Milo_myel_From_PBMC_compareDisease_cMon_1.svg"), width = 5.5, height = 3, bg='transparent')
ggsave(str_glue("{figurePath}Nhood_abundance_Milo_myel_From_PBMC_compareDisease_cMon_1.png"), width = 5.5, height = 3, bg='transparent')

#"#1f77b4" "#aec7e8" "#ff7f0e" "#ffbb78" "#2ca02c" "#98df8a" "#d62728" "#ff9896" "#9467bd" "#c5b0d5" "#8c564b" "#c49c94" "#e377c2" "#f7b6d2" "#7f7f7f" "#c7c7c7" "#bcbd22" "#dbdb8d" "#17becf" "#9edae5"

```

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=4}


# "Activated cMon"    "CCR7_CDE3" "cMon"   "Mon_SIRPA"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="Activated cMon",], disease_column = "disease_short", legend_title="Disease", axis.size = 15,
                             xlab = "", ylab = "Activated cMon Cells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_cMon_1.svg"), width = 5, height = 4, bg='transparent')
```

##### cMon

```{r fig.width=10, fig.height=6}
#[my_obs$dataset_id %in% c("GSE157278_Hong_2020", "GSE163314_Lefferts_2021", "Perez_2022") &

my_obs$Cell_type_scANVI = as.factor(my_obs$Cell_type_scANVI)

my_obs[my_obs$Cell_type_scANVI == "cMon", ] %>%
  dplyr::select(nhoo_test_condition,
                dataset_id,
                disease_short,
                Cell_type_scANVI) %>% mutate(
                  dataset_id = str_remove(dataset_id, "GSE[0-9]*_{1}"),
                  disease_short = str_remove(disease_short, "all_")
                ) %>% 
  group_by(nhoo_test_condition, dataset_id, Cell_type_scANVI, disease_short) %>% mutate(count = n()) %>% unique() %>%
  ggplot(aes(fill=nhoo_test_condition, y=count, x= reorder(disease_short, count))) + 
  geom_bar(position="stack", stat="identity", color = "white") + 
  labs(x = "", y = "Nhood Composition", fill = "nhood") +
  facet_wrap(~ dataset_id, scales = "free", ncol = 5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_bw() + #coord_flip() +
  scale_fill_manual(values = c(Enriched = "#459395", Depleted = "#FDA638", Non_sig = "grey50")) + #459395 EB7C69 FDA638
  scale_x_reordered() +
  theme(
    axis.text.x = element_text(
      hjust = 1,
      vjust = 0.5,
      size = 12,
      angle = 90
    ),
    axis.text.y = element_text(size = 12, color = "black"),
    #axis.text.x.bottom = element_blank(),
    legend.key.size = unit(0.6, 'cm'),
    legend.text = element_text(size = 10),
    #axis.line = element_line(colour = "black", linewidth = 0.6),
    legend.key.width = unit(0.4, "cm"),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    legend.key.height = unit(0.4, "cm")
  )

ggsave(str_glue("{figurePath}Nhood_abundance_Milo_myel_From_PBMC_compareDisease_cMon_2.svg"), width = 10, height = 6, bg='transparent')
ggsave(str_glue("{figurePath}Nhood_abundance_Milo_myel_From_PBMC_compareDisease_cMon_2.png"), width = 10, height = 6, bg='transparent')

#"#1f77b4" "#aec7e8" "#ff7f0e" "#ffbb78" "#2ca02c" "#98df8a" "#d62728" "#ff9896" "#9467bd" "#c5b0d5" "#8c564b" "#c49c94" "#e377c2" "#f7b6d2" "#7f7f7f" "#c7c7c7" "#bcbd22" "#dbdb8d" "#17becf" "#9edae5"

```

```{r fig.fullwidth = TRUE, fig.width=5, fig.height=4}

# "CCR7_CDE3" "cMon" "Mon_SIRPA"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="cMon",], disease_column = "disease_short", legend_title="Disease",
                             xlab = "", ylab = "cMon Cells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_cMon_2.svg"), width = 5, height = 4, bg='transparent')

```

##### Mon_SIRPA

```{r fig.fullwidth = TRUE, fig.width=6, fig.height=4}

# "CCR7_CDE3" "Mon_SIRPA"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="Mon_SIRPA",], disease_column = "disease_short", legend_title="Disease",
                             xlab = "", ylab = "Mon_SIRPA \nCells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_Mon_SIRPA.svg"), width = 6, height = 4, bg='transparent')

```

##### CCR7_CDE3

```{r fig.fullwidth = TRUE, fig.width=6, fig.height=4}

# "Activated cMon"    "CCR7_CDE3" "cMon"    "DC2"       "Non-cMon"  "Mon_SIRPA" "DC1"       "pDC"
plot_Cell_per_1000_per_donor(data = my_obs_0_2[my_obs_0_2$Cell_type_scANVI=="CCR7_CDE3",], disease_column = "disease_short", legend_title="Disease",
                             xlab = "", ylab = "CCR7_CDE3 \nCells count/1000 cells", color_dis_tis = color_disease_pbmc)



ggsave(str_glue("{figurePath}Cell_per_1000_per_donor_Milo_myel_From_PBMC_compareDisease_CCR7_CDE3.svg"), width = 6, height = 4, bg='transparent')

```

### Disease gene program

```{r}
colnames(colData(test_adata_))

test_adata_$Cell_type_scANVI %>% unique()

test_adata_$disease_short %>% unique()
```

#### Activated cMon

```{r}

levels(test_adata_$Cell_type_scANVI)
```

```{r}
test_adata_$nhoo = case_when(test_adata_$nhoo_test_condition=="Enriched" ~ "Enriched",
                                            TRUE ~ "Rest") %>% as.factor()

table(test_adata_$disease_short, test_adata_$dataset_id)


test_adata_$disease_nhoo = paste0(test_adata_$disease_short, "_", test_adata_$nhoo_test_condition) %>% as.factor()

table(test_adata_$disease_nhoo, test_adata_$Cell_type_scANVI)

```

```{r}
test_adata_$disease_short %>% unique()
```

```{r}
test_adata_$Cell_type_scANVI %>% unique()

test_adata_$Cell_type_scANVI = case_when(
  test_adata_$Cell_type_scANVI == "cMon_1" ~ "Activated cMon",
  test_adata_$Cell_type_scANVI == "cMon_2" ~ "cMon",
  TRUE ~ test_adata_$Cell_type_scANVI
)

test_adata_$Cell_type_scANVI %>% unique()
```

##### ALL disease

```{r}


cMon_1_deg = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset=T, hvgs = TRUE)



# cMon_1_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_,
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Activated cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset=T, hvgs = TRUE)


cMon_1_deg_dis_ctrl = run_deg(
  test_adata = test_adata_,#[,test_adata_$dataset_id=="GSE163314_Lefferts_2021"],
  lfc = 1,
  p.val = 0.05,
  cel_type = "Activated cMon",
  test_col="disease_short",
  ctrl="Healthy",
  var_column = "disease_short",
  var.included = c("all_CD", "all_SLE", "all_hashimoto", "all_sjo_syndr", "all_spondoarthritis", "Healthy"),
  covar = "dataset_id",
  covariable.dataset = T
)


write.csv(cMon_1_deg_dis_ctrl,
          str_glue("{degPathEnriched}cMon_1_deg_dis_ctrl.csv"))

cMon_1_deg_dis_ctrl_up = data.frame(name=cMon_1_deg_dis_ctrl$name[cMon_1_deg_dis_ctrl$dirFC=="Up"], celltype="Activated cMon")
write.csv(cMon_1_deg_dis_ctrl_up,
          str_glue("{degPathEnriched}Shared/cMon_1_deg_dis_ctrl.csv"))


cMon_1_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset=T, hvgs = TRUE)

```

##### SLE

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

```{r}
cMon_1_SLE_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)





# cMon_1_SLE_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Activated cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)




cMon_1_SLE_deg_dis_ctrl = run_deg(
  test_adata = test_adata_[,test_adata_$dataset_id=="Perez_2022"],
  lfc = 1,
  p.val = 0.05,
  cel_type = "Activated cMon",
  test_col="disease_short",
  ctrl="Healthy",
  var_column = "disease_short",
  var.included = c("Healthy", "all_SLE"),
  covariable.dataset = F
)


write.csv(cMon_1_SLE_deg_dis_ctrl,
          str_glue("{degPathEnriched}cMon_1_SLE_deg_dis_ctrl.csv"))




cMon_1_SLE_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Sjo_syndr

```{r}
cMon_1_Sjo_syndr_deg = disease_nhoo_program(
  test_adata = test_adata_[, test_adata_$dataset_id == "GSE157278_Hong_2020"],
  
  lfc = 1,
  
  dir_fc = "Enriched",
  
  p.val = 0.05,
  
  cel_type = "Activated cMon",
  var_column = "disease_short",
  comparison = "selected_nhoo",
  covariable.dataset = FALSE,
  hvgs = TRUE
)



# cMon_1_Sjo_syndr_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Activated cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)


cMon_1_Sjo_syndr_deg_dis_ctrl = run_deg(
  test_adata = test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
  lfc = 1,
  p.val = 0.05,
  cel_type = "Activated cMon",
  test_col="disease_short",
  ctrl="Healthy",
  var_column = "disease_short",
  var.included = c("Healthy", "all_sjo_syndr"),
  covariable.dataset = F
)


write.csv(cMon_1_Sjo_syndr_deg_dis_ctrl,
          str_glue("{degPathEnriched}cMon_1_Sjo_syndr_deg_dis_ctrl.csv"))


cMon_1_Sjo_syndr_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Spondoarthritis

```{r}
cMon_1_Spondoarthritis_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



# cMon_1_Spondoarthritis_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Activated cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)


cMon_1_Spondoarthritis_deg_dis_ctrl = run_deg(
  test_adata = test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021"],
  lfc = 1,
  p.val = 0.05,
  cel_type = "Activated cMon",
  test_col="disease_short",
  ctrl="Healthy",
  var_column = "disease_short",
  var.included = c("all_spondoarthritis", "Healthy"),
  covariable.dataset = F
)


write.csv(cMon_1_Spondoarthritis_deg_dis_ctrl,
          str_glue("{degPathEnriched}cMon_1_Spondoarthritis_deg_dis_ctrl.csv"))


cMon_1_Spondoarthritis_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### CD

```{r}
cMon_1_CD_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = F, hvgs = TRUE)



# cMon_1_CD_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Activated cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = F, hvgs = TRUE)


cMon_1_CD_deg_dis_ctrl = run_deg(
  test_adata = test_adata_,#[,test_adata_$dataset_id=="GSE163314_Lefferts_2021"],
  lfc = 1,
  p.val = 0.05,
  cel_type = "Activated cMon",
  test_col="disease_short",
  ctrl="Healthy",
  var_column = "disease_short",
  var.included = c("all_CD", "Healthy"),
  covar = "dataset_id",
  covariable.dataset = T
)


write.csv(cMon_1_CD_deg_dis_ctrl,
          str_glue("{degPathEnriched}cMon_1_CD_deg_dis_ctrl.csv"))


cMon_1_CD_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & 
                                                                       
                                                                       test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = F, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Hashimoto

```{r}
cMon_1_hashimoto_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_1_hashimoto_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],

                                lfc = 1,

                                dir_fc = "Enriched",

                                p.val = 0.05,

                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


# cMon_1_hashimoto_deg_dis_ctrl = run_deg(
#   test_adata = test_adata_,#[,test_adata_$dataset_id=="GSE163314_Lefferts_2021"],
#   lfc = 1,
#   p.val = 0.05,
#   cel_type = "Activated cMon",
#   test_col="disease_short",
#   ctrl="Healthy",
#   var_column = "disease_short",
#   var.included = c("all_hashimoto", "Healthy"),
#   covar = "dataset_id",
#   covariable.dataset = T
# )


# write.csv(
#   cMon_1_hashimoto_deg_dis_ctrl,
#   str_glue("{degPathEnriched}cMon_1_hashimoto_deg_dis_ctrl.csv")
# )

cMon_1_hashimoto_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

#### cMon

```{r}
test_adata_$nhoo = case_when(test_adata_$nhoo_test_condition=="Depleted" ~ "Depleted",
                                            TRUE ~ "Rest") %>% as.factor()

table(test_adata_$disease_short, test_adata_$nhoo)
```

```{r}

levels(test_adata_$Cell_type_scANVI)
```

```{r}
test_adata_$nhoo = case_when(test_adata_$nhoo_test_condition=="Enriched" ~ "Enriched",
                                            TRUE ~ "Rest") %>% as.factor()

table(test_adata_$disease_short, test_adata_$dataset_id)

```

```{r}
test_adata_$nhoo %>% unique()
```

```{r}

cMon_2_deg = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo", hvgs = TRUE)



cMon_2_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl", hvgs = TRUE)


cMon_2_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo", hvgs = TRUE)


```

##### SLE

```{r}
cMon_2_SLE_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_2_SLE_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_2_SLE_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Sjo_syndr

```{r}
cMon_2_Sjo_syndr_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_2_Sjo_syndr_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_2_Sjo_syndr_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Spondoarthritis

```{r}
# cMon_2_Spondoarthritis_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Depleted",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="cMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# 
# cMon_2_Spondoarthritis_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Depleted",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="cMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# cMon_2_Spondoarthritis_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Depleted",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="cMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### CD

```{r}
cMon_2_CD_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = F, hvgs = TRUE)



cMon_2_CD_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = F, hvgs = TRUE)


cMon_2_CD_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & 
                                                                       
                                                                       test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = F, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Hashimoto

```{r}
cMon_2_hashimoto_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_2_hashimoto_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_2_hashimoto_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

#### Activated cMon and cMon together

```{r}

cMon_deg = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo", hvgs = TRUE)



cMon_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl", hvgs = TRUE)


cMon_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo", hvgs = TRUE)

```

##### SLE

```{r}
cMon_SLE_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_SLE_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_SLE_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Sjo_syndr

```{r}
cMon_Sjo_syndr_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_Sjo_syndr_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_Sjo_syndr_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Spondoarthritis

```{r}
cMon_Spondoarthritis_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_Spondoarthritis_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_Spondoarthritis_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### CD

```{r}
cMon_CD_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = F, hvgs = TRUE)



cMon_CD_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = F, hvgs = TRUE)


cMon_CD_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & 
                                                                       
                                                                       test_adata_$disease_short %in% c("all_CD", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = F, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Hashimoto

```{r}
cMon_hashimoto_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="selected_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)



cMon_hashimoto_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl",
                                covariable.dataset = FALSE, hvgs = TRUE)


cMon_hashimoto_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type=c("Activated cMon", "cMon"),
                                var_column = "disease_short",
                                comparison="between_nhoo",
                                covariable.dataset = FALSE, hvgs = TRUE)

```

#### Non_CMon

```{r}

levels(test_adata_$Cell_type_scANVI)
```

```{r}
test_adata_$nhoo = case_when(test_adata_$nhoo_test_condition=="Enriched" ~ "Enriched",
                                            TRUE ~ "Rest") %>% as.factor()

table(test_adata_$disease_short, test_adata_$dataset_id)


test_adata_$disease_nhoo = paste0(test_adata_$disease_short, "_", test_adata_$nhoo_test_condition) %>% as.factor()

table(test_adata_$disease_nhoo, test_adata_$Cell_type_scANVI)

```

```{r}
test_adata_$dataset_id %>% unique()
```

```{r}


Non_cMon_enriched_deg = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Non_cMon",
                                var_column = "disease_short",
                                comparison="selected_nhoo", hvgs = TRUE)



Non_cMon_enriched_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Non_cMon",
                                var_column = "disease_short",
                                comparison="dis_vs_ctrl", hvgs = TRUE)


Non_cMon_enriched_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Non_cMon",
                                var_column = "disease_short",
                                comparison="between_nhoo", hvgs = TRUE)

```

##### SLE

```{r}

# Non_cMon_SLE_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# 
# Non_cMon_SLE_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# Non_cMon_SLE_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="Perez_2022"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Sjo_syndr

```{r}
# Non_cMon_Sjo_syndr_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# 
# Non_cMon_Sjo_syndr_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# Non_cMon_Sjo_syndr_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE157278_Hong_2020"],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Spondoarthritis

```{r}
# Non_cMon_Spondoarthritis_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# 
# Non_cMon_Spondoarthritis_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# Non_cMon_Spondoarthritis_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_spondoarthritis", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### CD

```{r}
# Non_cMon_CD_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = F, hvgs = TRUE)
# 
# 
# 
# Non_cMon_CD_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & test_adata_$disease_short %in% c("all_CD", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = F, hvgs = TRUE)
# 
# 
# Non_cMon_CD_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$dataset_id=="GSE163314_Lefferts_2021" & 
#                                                                        
#                                                                        test_adata_$disease_short %in% c("all_CD", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = F, hvgs = TRUE)

```

```{r}
table(test_adata_$disease_short, test_adata_$dataset_id)
```

##### Hashimoto

```{r}
# Non_cMon_hashimoto_deg = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="selected_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# 
# Non_cMon_hashimoto_deg_dis_ctrl = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="dis_vs_ctrl",
#                                 covariable.dataset = FALSE, hvgs = TRUE)
# 
# 
# Non_cMon_hashimoto_deg_dis_nhoo = disease_nhoo_program(test_adata=test_adata_[,test_adata_$disease_short %in% c("all_hashimoto", "Healthy")],
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non_CMon",
#                                 var_column = "disease_short",
#                                 comparison="between_nhoo",
#                                 covariable.dataset = FALSE, hvgs = TRUE)

```

#### enriched_vs_depleted

##### ALL disease

```{r}


cMon_1_enriched_disease = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Enriched",
                                
                                p.val = 0.05,
                                
                                cel_type="Activated cMon",
                                var_column = "disease_short",
                                comparison = "enriched_vs_depleted",
                                covariable.dataset=T, hvgs = T)


cMon_2_depleted_disease = disease_nhoo_program(test_adata=test_adata_,
                                
                                lfc = 1,
                                
                                dir_fc = "Depleted",
                                
                                p.val = 0.05,
                                
                                cel_type="cMon",
                                var_column = "disease_short",
                                comparison = "enriched_vs_depleted",
                                covariable.dataset=T, hvgs = T)


# Non_cMon_enriched_disease = disease_nhoo_program(test_adata=test_adata_,
#                                 
#                                 lfc = 1,
#                                 
#                                 dir_fc = "Enriched",
#                                 
#                                 p.val = 0.05,
#                                 
#                                 cel_type="Non-CMon",
#                                 var_column = "disease_short",
#                                 comparison = "enriched_vs_depleted",
#                                 covariable.dataset=T, hvgs = T)



```

```{r}
cMon_1_enriched_disease %>% filter(abs(lfc) < 20, abs(lfc_dis_enriched) < 20) %>% 
  ggplot(aes(x=lfc, y=lfc_dis_enriched)) +
  geom_point()
```

```{r}
cor.test(cMon_1_enriched_disease$lfc, cMon_1_enriched_disease$lfc_dis_enriched)
```

```{r}
cMon_1_enriched_disease = cMon_1_enriched_disease %>% filter(dirFC_dis_enriched != "NS") %>% filter(dirFC_rest_dis_depleted != dirFC_dis_enriched) %>% arrange(desc(lfc_dis_enriched))
```

```{r}
cMon_2_depleted_disease %>% filter(abs(lfc) < 20, abs(lfc_dis_enriched) < 20) %>% 
  ggplot(aes(x=lfc, y=lfc_dis_enriched)) +
  geom_point()
```

```{r}
cor.test(cMon_2_depleted_disease$lfc, cMon_2_depleted_disease$lfc_dis_enriched)
```

```{r}

cMon_2_depleted_disease %>% filter(abs(lfc) < 20, abs(lfc_dis_enriched) < 20) %>% 
  ggplot(aes(x=lfc, y=lfc_dis_enriched)) +
  geom_point()
```

### Comparison c_Mon_1 vs c_Mon_2 (only in disease)

```{r}

table(test_adata_$Cell_type_scANVI, test_adata_$disease_short)
```

```{r}
test_adata_$dataset_disease = paste0(test_adata_$dataset_id, "_", test_adata_$disease_short) %>% as.factor()

#table(test_adata_$dataset_disease)
```

```{r}
test_adata_$cell_disease_nhoo = paste0(test_adata_$Cell_type_scANVI, "_", test_adata_$disease_nhoo) %>% as.factor()

table(test_adata_$cell_disease_nhoo)
```

#### All

```{r}

test_adata_$disease_bin = ifelse(test_adata_$disease_short=="Healthy", "Healthy", "Disease") %>% factor(levels = c("Healthy", "Disease"))
  
  
test_adata_$Cell_disease_bin = paste0(test_adata_$Cell_type_scANVI, "_", test_adata_$disease_bin)

All_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_disease_bin",
  ctrl=c("cMon_1_Healthy", "cMon_2_Disease", "cMon_2_Healthy"),
  var_column = "disease_short",
  var.included = c("all_CD", "all_SLE", "all_hashimoto", "all_sjo_syndr", "all_spondoarthritis", "Healthy"),
  covariable.dataset = TRUE,
  covar = "dataset_disease"
)


All_cMon_2_vs_cMon_1 = All_cMon_2_vs_cMon_1 %>% arrange(desc(AUC))

```

```{r}

All_cMon_2_vs_cMon_1_healthy = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("Healthy"),
  covariable.dataset = TRUE,
  covar = "dataset_id"
)

All_cMon_2_vs_cMon_1_disease = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_CD", "all_SLE", "all_hashimoto", "all_sjo_syndr", "all_spondoarthritis"),
  covariable.dataset = T,
  covar = "dataset_id"
)


cMon_1_vs_cMon_2 = setdiff(paste0(All_cMon_2_vs_cMon_1_disease$name[All_cMon_2_vs_cMon_1_disease$AUC>=0.7], "_", All_cMon_2_vs_cMon_1_disease$dirFC[All_cMon_2_vs_cMon_1_disease$AUC>=0.7]),
        paste0(All_cMon_2_vs_cMon_1_healthy$name[All_cMon_2_vs_cMon_1_healthy$AUC>=0.7], "_", All_cMon_2_vs_cMon_1_healthy$dirFC[All_cMon_2_vs_cMon_1_healthy$AUC>=0.7])) %>% str_remove(., "_.*")


cMon_1_vs_cMon_2

```

```{r fig.width=15, fig.height=7}
levels(test_adata_$Cell_type_scANVI)
levels(test_adata_$disease_short)
All_cMon_2_vs_cMon_1_disease = All_cMon_2_vs_cMon_1_disease %>% arrange(desc(AUC))

test_adata = test_adata_[, test_adata_$Cell_type_scANVI %in% c("Activated cMon", "cMon") & test_adata_$disease_short == "all_CD"]

#test_adata$disease_bin = ifelse(test_adata$disease_short == "Remission_RA", "Remission_RA", "Naïve_treat_RA") %>% as.factor()


plot_violin_facet(
  sce = test_adata,
  gene = c("ICAM2", "CD44", "ITGAL", "ITGAM", "CCR1", "CCR2", "S100A8", "S100A9"),
  xlab = "",
  ylab = "",
  size = 3,
  shape = 20,
  cell_type = c("Activated cMon", "cMon"),
  gg_x = "Cell_type_scANVI",
  angle = 90,
  colmap = vega_20,
  base_size = 14,
  base_line_size = 0.5,
  base_family = "sans",
  base_fontface = "bold",
  legend.title = "Groups",
  strip.colour = "transparent",
  strip.fill = "transparent",
  plot.title.position = "plot",
  plot.title = "",
  legend.position = "right",
  ncol = 6,
  show.x.text = FALSE,
  figure_Path,
  fig.width = 8,
  fig.height = 3,
  save = F,
  stat = T,
  stat.ref = "cMon",
  stat.method = "t.test",
  stat.paired = FALSE,
  label.y = NULL,
  violin = T,
  scale.data = F,
  include_violin_boxplot = T,
  violin_boxplot_w = 0.1
)


```

##### GSEA

```{r}
All_cMon_2_vs_cMon_1_2 = All_cMon_2_vs_cMon_1_disease[,c(1,7,9)] %>% arrange(desc(AUC))

#colnames(All_cMon_2_vs_cMon_1_2) = NULL

#rownames(All_cMon_2_vs_cMon_1_2) = NULL

write.csv(x = as.data.frame(All_cMon_2_vs_cMon_1_2)[,1:2], file = "G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/All_cMon_2_vs_cMon_1.tsv",
          row.names = F)
```

```{r}
table(test_adata_$dataset_id, test_adata_$disease_short)
```

#### CD

```{r}

IBD_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_CD"),
  covariable.dataset = F,
  covar = "dataset_id"
)


```

#### SLE

```{r}

SLE_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_SLE"),
  covariable.dataset = F,
  covar = "dataset_id"
)


```

#### hashimoto

```{r}

Hashimoto_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_hashimoto"),
  covariable.dataset = F,
  covar = "dataset_id"
)


```

#### Sjo_syndr

```{r}

sjo_syndr_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_sjo_syndr"),
  covariable.dataset = F,
  covar = "dataset_id"
)


```

#### all_spondoarthritis

```{r}

spondoarthritis_cMon_2_vs_cMon_1 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("cMon", "Activated cMon"),
  test_col="Cell_type_scANVI",
  ctrl=c("cMon"),
  var_column = "disease_short",
  var.included = c("all_spondoarthritis"),
  covariable.dataset = F,
  covar = "dataset_id"
)


```

##### Common cMon1_vs_cMon2

```{r}
rownames(All_cMon_2_vs_cMon_1) = paste0(All_cMon_2_vs_cMon_1$name, "_", All_cMon_2_vs_cMon_1$dirFC)

rownames(IBD_cMon_2_vs_cMon_1) = paste0(IBD_cMon_2_vs_cMon_1$name, "_", IBD_cMon_2_vs_cMon_1$dirFC)

rownames(Hashimoto_cMon_2_vs_cMon_1) = paste0(Hashimoto_cMon_2_vs_cMon_1$name, "_", Hashimoto_cMon_2_vs_cMon_1$dirFC)

rownames(sjo_syndr_cMon_2_vs_cMon_1) = paste0(sjo_syndr_cMon_2_vs_cMon_1$name, "_", sjo_syndr_cMon_2_vs_cMon_1$dirFC)


rownames(spondoarthritis_cMon_2_vs_cMon_1) = paste0(spondoarthritis_cMon_2_vs_cMon_1$name, "_", spondoarthritis_cMon_2_vs_cMon_1$dirFC)

rownames(SLE_cMon_2_vs_cMon_1) = paste0(SLE_cMon_2_vs_cMon_1$name, "_", SLE_cMon_2_vs_cMon_1$dirFC)

#intersect of DEG

common_deg_cMon_1_vs_cMon_2 = base::Reduce(intersect, list(rownames(IBD_cMon_2_vs_cMon_1),
                             rownames(Hashimoto_cMon_2_vs_cMon_1),
                             rownames(sjo_syndr_cMon_2_vs_cMon_1),
                             rownames(spondoarthritis_cMon_2_vs_cMon_1),
                             rownames(SLE_cMon_2_vs_cMon_1)))

rownames(All_cMon_2_vs_cMon_1_disease) = paste0(All_cMon_2_vs_cMon_1_disease$name, "_", All_cMon_2_vs_cMon_1_disease$dirFC)

commoncMon_1_vs_cMon_2 = intersect(common_deg_cMon_1_vs_cMon_2, rownames(All_cMon_2_vs_cMon_1_disease)) %>% str_remove(., "_.*")


commoncMon_1_vs_cMon_2 = as.data.frame(All_cMon_2_vs_cMon_1_disease[All_cMon_2_vs_cMon_1_disease$name%in%commoncMon_1_vs_cMon_2,]) 
```

##### GSEA results

```{r}
gsea_KEGG_2021_Human = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/GSEA_CMon1_vs_cMon_2_KEGG_2021_Human.csv")
```

```{r}
gsea_GO_Biological_Process_2023 = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/GSEA_CMon1_vs_cMon_2_GO_Biological_Process_2023.csv")
```

```{r}
gsea_MSigDB_Hallmark_2020 = read.csv("C:/Users/bidos/OneDrive/Posdoc/bepe/Results_Sanger/sanger_gdrive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/GSEA_CMon1_vs_cMon_2_MSigDB_Hallmark_2020.csv")
```

```{r}
gsea_MSigDB_Reactome_2022 = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/GSEA_CMon1_vs_cMon_2_Reactome_2022.csv")
```

```{r}
gsea_MSigDB_WikiPathway_2021_Human = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/GSEA/GSEA_CMon1_vs_cMon_2_WikiPathway_2021_Human.csv")
```

##### Heatmap

```{r fig.width=11, fig.height=18}
#| fig-width: 25
#| fig-height: 100


cel_nh = levels(test_adata_$Cell_type_scANVI)


n_top = 25

test_adata_sub = test_adata_[, test_adata_$Cell_type_scANVI %in% c("Activated cMon", "cMon")]

#test_adata_sub = test_adata_[, test_adata_$Cell_type_scANVI %in% c("Intermediate_macrophage", "Mac_MERTK_hi_TREM2_low", "Mac_NUPR1_TREM2_hi", #"Mac_S100A8_prod", "Mo_Mac_CX3CR1", "Mo_Mac_LYVE1_plus")]

test_adata_prep = prep_de_data(depl_ = "", enrch_ = "", test_adata = test_adata_sub, nhood = FALSE, hvgs = FALSE, cel_of_int_vs_rest = T, cell_type = "Activated cMon")

#cel_cleaned = cel %% str_replace_all("_plus", "+")
#cel_cleaned = str_glue("Depleted {cel_cleaned}")

rownames(cMon_1_deg_dis_ctrl) = paste0(cMon_1_deg_dis_ctrl$name, "_", cMon_1_deg_dis_ctrl$dirFC)

rownames(cMon_2_deg_dis_ctrl) = paste0(cMon_2_deg_dis_ctrl$name, "_", cMon_2_deg_dis_ctrl$dirFC)


cMon1 = cMon_1_deg_dis_ctrl[rownames(cMon_1_deg_dis_ctrl) %in% rownames(cMon_2_CD_deg_dis_ctrl),]

test_adata_prep$disease_short = test_adata_prep$disease_short %>% 
  str_replace(., "CD", "IBD") %>% str_remove(., "all_") %>% 
  str_to_sentence() %>% str_replace(., "Sle", "SLE") %>% str_replace(., "Ibd", "IBD") %>% as.factor()
  
plot_heatmap_wout_bar(
  test_adata = test_adata_prep[,test_adata_prep$disease_short!="Healthy"],
  cel.selected = "Activated cMon",
  selected_gene = unique(c(str_split_1(gsea_MSigDB_Hallmark_2020$leading_edge[1], pattern = ","),
                    str_split_1(gsea_MSigDB_Hallmark_2020$leading_edge[2], pattern = ","))),
  n_top = 15,
  aggregate = TRUE,
  show_column_names = TRUE,
  aggr_fun = "mean",
  col_pal = vega_20,
  test.column = "Cell_type_scANVI",
  column.label = "disease_short",
  col_test = c("grey", "grey50"),
  scale = TRUE,
  remove_inf = TRUE,
  cluster.col = F,
  cluster.row = T,
  labels_annot_bar = TRUE,
  label.only.disease = FALSE,
  figure_Path = str_glue("{figurePath}Heatmap_cMon_1_vs_cMon_1_only_disease"),
  save = F,
  fig.width = 11,
  fig.height = 18,
  font = 35,
  title.font = 35,
  title.face = "bold",
  legend_h = 10,
  legend_w = 1,
  fontface = "bold",
  grid_width = 15,
  column_names_rot = 45
)

```

## Save

```{r}
save.image(file = "G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/PBMC_Myeloid_session.RData")
```

```{r}
load(file = "G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/PBMC_Myeloid_session.RData")
```

```{r}
fraction_myel = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/PBMC_Myeloid_individual_disease_Milo_adata_reajusted_Summary_stat_fraction.csv")

fraction_myel = tibble::column_to_rownames(fraction_myel, var = "X")


fraction_myel_meta = read.csv("G:/My Drive/Projects/SCVI_Results/Immune_compartment_PBMC/Subset_Agosto_2023/Milo/individual_disease_nhoo/PBMC_Myeloid_individual_disease_Milo_adata_reajusted_Summary_stat_metadata.csv")

fraction_myel_meta = tibble::column_to_rownames(fraction_myel_meta, var = "X")
```

```{r}
fraction_myel_2 = pivot_longer(fraction_myel, cols = 1:ncol(fraction_myel), names_to = "donor", values_to = "Percentage")
```

```{r}

counts(test_adata_) <- assay(test_adata_)
  
test_adata_ = logNormCounts(test_adata_)

test_adata_count = assay(test_adata_[, test_adata_$Cell_type_scANVI == "Activated cMon"], "logcounts")%>% t() %>% as.data.frame()

#cMon_1_fract = test_adata_count %>% summarise_at(-1, ~sum(.>=10)/n())%>% t()


cMon_1_fract = colSums(test_adata_count) %>% as.data.frame()
colnames(cMon_1_fract) = "Activated cMon"

test_adata_count = assay(test_adata_[, test_adata_$Cell_type_scANVI == "cMon"], "logcounts")%>% t() %>% as.data.frame()


#cMon_2_fract = test_adata_count %>% summarise_at(-1, ~sum(.>=10)/n()) %>% t()

cMon_2_fract = colSums(test_adata_count) %>% as.data.frame()

colnames(cMon_2_fract) = "cMon"

all.equal(rownames(cMon_2_fract), rownames(cMon_1_fract))

cMon_1_cMon_2_fract = cbind(as.data.frame(cMon_1_fract), as.data.frame(cMon_2_fract)) %>% 
  mutate(ratio = `Activated cMon`/cMon)


cMon_1_cMon_2_fract %>% ggplot(aes(x=`Activated cMon`, y=cMon)) +
  geom_point()

```

```{r}
plot_ggplot_gene_col(sce = test_adata_, gene="IL1B", title="IL1B", subtitle = "", xlab = "", ylab = "",
                                size = 2, shape = 20, cell_type = c("Activated cMon", "cMon"), gg_x="Cell_type_scANVI", angle = 90, colmap = vega_20,
                                base_size = 14, base_family = "sans", base_fontface = "bold", 
                                plot.title.position = "plot", legend.position = "right", show.x.text = FALSE)
```

### Venn diagram

```{r}
library(VennDiagram)
library(RColorBrewer)
```

### Activated cMon

```{r}
cMon_1_hashimoto_deg_dis_ctrl = read.csv(
  str_glue("{degPathEnriched}cMon_1_hashimoto_deg_dis_ctrl.csv")
)
# Generate 3 sets
set1 <- paste(cMon_1_Sjo_syndr_deg_dis_ctrl$name, cMon_1_Sjo_syndr_deg_dis_ctrl$dirFC, sep = "_")
set2 <- paste(cMon_1_hashimoto_deg_dis_ctrl$name, cMon_1_hashimoto_deg_dis_ctrl$dirFC, sep = "_")
set3 <- paste(cMon_1_SLE_deg_dis_ctrl$name, cMon_1_SLE_deg_dis_ctrl$dirFC, sep = "_")

# Prepare a palette of 3 colors with R colorbrewer:

myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(
        x = list(set1, set2, set3),
        category.names = c("Sjo_syndr" , "hashimoto" , "SLE"),
        filename = str_glue('{figurePath}cMon_1_venn_diagramm.png'),
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
        # Set names
        cat.cex = 0.6,
        #cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135),
        cat.dist = c(0.055, 0.055, 0.085),
        cat.fontfamily = "sans",
        rotation = 1
)

```

```{r}
intersect(intersect(set1, set2), set3)
```

#### Heatmap

```{r}
selected_gene = gsub("_.*$", "", intersect(intersect(set1, set2), set3))


```

```{r fig.width=15, fig.height=7.5}
#| fig-width: 25
#| fig-height: 100
cel_nh = levels(test_adata_$Cell_type_scANVI)



test_adata_sub = test_adata_[, test_adata_$Cell_type_scANVI=="Activated cMon" & test_adata_$disease_short %in% c("Healthy", "all_SLE", "all_hashimoto", "all_sjo_syndr")]



test_adata_prep = prep_de_data(depl_ = "", enrch_ = "", test_adata = test_adata_sub, nhood = FALSE, hvgs = FALSE, cel_of_int_vs_rest = T, cell_type = "Activated cMon")

test_adata_prep$disease_short = str_remove(test_adata_prep$disease_short, "all_") %>% as.factor()

test_adata_prep$disease_dataset = paste(test_adata_prep$disease_short, test_adata_prep$dataset_id, "_") %>% as.factor()

plot_heatmap(
  test_adata = test_adata_prep,
  selected_gene = selected_gene,
  cel.selected = "Activated cMon",
  #cel_cleaned,
  aggregate = T,
  n_top = 20,
  show_column_names = T,
  col_pal = vega_20,
  test.column = "disease_dataset",
  #
  column.label2 = "disease_short",
  labels_annot_bar = F,
  label.only.disease = TRUE,
  scale = T,
  remove_inf = F,
  cluster.col = F,
  figure_Path = str_glue("{figurePath}Heatmap_cMon_1_shared_disease_DE_genes"),
  save = F,
  fig.width=7,
  fig.height=8.5,
)



```

### cMon

```{r}
# Generate 3 sets of 200 words
set1 <- paste(cMon_2_Sjo_syndr_deg_dis_ctrl$name, cMon_2_Sjo_syndr_deg_dis_ctrl$dirFC, sep = "_")
set2 <- paste(cMon_2_hashimoto_deg_dis_ctrl$name, cMon_2_hashimoto_deg_dis_ctrl$dirFC, sep = "_")
set3 <- paste(cMon_2_SLE_deg_dis_ctrl$name, cMon_2_SLE_deg_dis_ctrl$dirFC, sep = "_")

# Prepare a palette of 3 colors with R colorbrewer:

myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(
        x = list(set1, set2, set3),
        category.names = c("Sjo_syndr" , "hashimoto" , "SLE"),
        filename = str_glue('{figurePath}cMon_2_venn_diagramm.png'),
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
        # Set names
        cat.cex = 0.6,
        #cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135),
        cat.dist = c(0.055, 0.055, 0.085),
        cat.fontfamily = "sans",
        rotation = 1
)

```

```{r}
intersect(intersect(set1, set2), set3)
```

### Non-CMon

```{r}
# Generate 3 sets of 200 words
set1 <- paste(cMon_1_Sjo_syndr_deg_dis_ctrl$name, cMon_1_Sjo_syndr_deg_dis_ctrl$dirFC, sep = "_")
set2 <- paste(cMon_1_hashimoto_deg_dis_ctrl$name, cMon_1_hashimoto_deg_dis_ctrl$dirFC, sep = "_")
set3 <- paste(cMon_1_SLE_deg_dis_ctrl$name, cMon_1_SLE_deg_dis_ctrl$dirFC, sep = "_")

# Prepare a palette of 3 colors with R colorbrewer:

myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(
        x = list(set1, set2, set3),
        category.names = c("Sjo_syndr" , "hashimoto" , "SLE"),
        filename = str_glue('{figurePath}cMon_1_venn_diagramm.png'),
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",
        
        # Set names
        cat.cex = 0.6,
        #cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135),
        cat.dist = c(0.055, 0.055, 0.085),
        cat.fontfamily = "sans",
        rotation = 1
)

```

```{r}
intersect(intersect(set1, set2), set3)
```

```{r}
levels(test_adata_$disease_short)

cMon_1_vs_cMon_2 = run_deg(
  test_adata = test_adata_,
  lfc = 1,
  p.val = 0.05,
  cel_type = c("Mac_S100A8_prod"),
  test_col="disease_short",
  ctrl=c("Remission_RA"),
  var_column = "disease_short",
  var.included = c("Remission_RA", "Resitant_treat_RA"),
  covariable.dataset = F,
  covar = "dataset_id"
)

Mac_S100A8_Remission_vs_Resistant = Mac_S100A8_Remission_vs_Resistant %>% arrange(desc(AUC))
```

### Egress markers

```{r fig.width=20, fig.height=5}
levels(test_adata_$Cell_type_scANVI)

#test_adata = test_adata_[, test_adata_$disease_short %in% c("Naïve_treat_RA", "Resitant_treat_RA")]



plot_violin_facet(
  sce = test_adata,
  gene = Mac_S100A8_Resistant_vs_Naive_treat_RA$name[1:20],
  xlab = "",
  ylab = "",
  size = 3,
  shape = 20,
  cell_type = c("Activated cMon", "cMon"),
  gg_x = "disease_short",
  angle = 90,
  colmap = color_disease_tis,
  base_size = 14,
  base_line_size = 0.5,
  base_family = "sans",
  base_fontface = "bold",
  legend.title = "Groups",
  strip.colour = "black",
  strip.fill = "grey80",
  plot.title.position = "plot",
  plot.title = "S100A8+ macrophages DEG (Resitant_treat_RA vs Naïve_treat_RA)",
  legend.position = "right",
  ncol = 10,
  show.x.text = FALSE,
  figure_Path = "",
  fig.width = 8,
  fig.height = 3,
  save = T,
  stat = F,
  stat.ref = "Remission_RA",
  stat.method = "t.test",
  stat.paired = FALSE,
  label.y = NULL,
  violin = F,
  include_violin_boxplot = T,
  violin_boxplot_w = 0.1
)


```


### Figure 3D

```{r fig.width=11, fig.height=18}
#| fig-width: 25
#| fig-height: 100


cel_nh = unique(test_adata_$Cell_type_scANVI)


n_top = 25

test_adata_sub = test_adata_[, test_adata_$Cell_type_scANVI %in% c("Activated cMon", "cMon")]

test_adata_sub$disease_short = str_remove(test_adata_sub$disease_short, "all_")

test_adata_sub$disease_short = str_replace(test_adata_sub$disease_short, "CD", "IBD")

test_adata_sub$disease_short = str_replace(test_adata_sub$disease_short, "hashimoto", "Hashimoto's thyroiditis")

test_adata_sub$disease_short = str_replace(test_adata_sub$disease_short, "spondoarthritis", "Spondyloarthritis")

test_adata_sub$disease_short = str_replace(test_adata_sub$disease_short, "sjo_syndr", "Sjögren Syndrome") 

#test_adata_sub = test_adata_[, test_adata_$Cell_type_scANVI %in% c("Intermediate_macrophage", "Mac_MERTK_hi_TREM2_low", "Mac_NUPR1_TREM2_hi", #"Mac_S100A8_prod", "Mo_Mac_CX3CR1", "Mo_Mac_LYVE1_plus")]

test_adata_prep = prep_de_data(depl_ = "", enrch_ = "", test_adata = test_adata_sub, nhood = FALSE, hvgs = FALSE, cel_of_int_vs_rest = T, cell_type = "Activated cMon")

test_adata_prep$Cell_type_scANVI %>% unique()
test_adata_prep$disease_short %>% unique()

test_adata_prep$disease_short = factor(test_adata_prep$disease_short, levels = c("Healthy", sort(as.character(
  unique(test_adata_prep$disease_short[!(test_adata_prep$disease_short %in% c("Healthy"))])
))))
  
test_adata_prep$disease_short %>% unique()

gsea_net_heat = rbind(
  data.frame(
    Term = gsea_MSigDB_Hallmark_2020$Term[1],
    symbol = str_split_1(gsea_MSigDB_Hallmark_2020$leading_edge[1], pattern = ",")
  ),
  data.frame(
    Term = gsea_MSigDB_Hallmark_2020$Term[2],
    symbol = str_split_1(gsea_MSigDB_Hallmark_2020$leading_edge[2], pattern = ",")
  )
)


gsea_net_heat$value = 1

gsea_net_heat = pivot_wider(gsea_net_heat, names_from = Term, values_from = value, values_fill = 0)

gsea_net_heat = tibble::column_to_rownames(gsea_net_heat, var="symbol")

colnames(gsea_net_heat) = str_replace(colnames(gsea_net_heat), "Response", "\nResponse")

plot_heatmap_wout_bar(
  test_adata = test_adata_prep[,test_adata_prep$disease_short!="Healthy"],
  cel.selected = "Activated cMon",
  selected_gene = rownames(gsea_net_heat),
  gsea_net_heat = gsea_net_heat,
  ref.vs.other = FALSE,
  n_top = 15,
  aggregate = TRUE,
  show_column_names = TRUE,
  aggr_fun = "mean",
  col_pal = vega_20,
  test.column = "Cell_type_scANVI",
  column.label = "disease_short",
  col_test = c("grey", "grey50"),
  scale = TRUE,
  remove_inf = TRUE,
  cluster.col = FALSE,
  labels_annot_bar = TRUE,
  label.only.disease = FALSE,
  figure_Path = str_glue("{figurePath}Heatmap_cMon_1_vs_cMon_1_only_disease"),
  save = F,
  fig.width = 11,
  fig.height = 18,
  font = 25,
  path.col.font = 20,
  title.font = 18,
  title.face = "bold",
  fontface = "bold",
  grid_width = 15,
  legend_h = 10,
  legend_w = 2,
  grid_width1 = 2,
  grid_height1 = 2,
  title_g = 10, broadcell = "macrophages",
  grid_width_row_annot = 8, 
  rowannot = T, path_title = "Pathways",
  column.name.col = "grey20"
)


```
